---
title: "Monitoring transmission intensity of trachoma with serology"
subtitle: "Create analysis datasets"
author: "Contributors: Christine Tedijanto christine.tedijanto@ucsf.edu, Ben Arnold ben.arnold@ucsf.edu"
date: "Last updated: `r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
number_sections: false
---

**Objective:** Read in the overall harmonized trachoma serology datasets and filter them to surveys and measurements that meet the analysis inclusion criteria:

* For studies with multiple years of data, focus on the most recent year (and in the case of studies with PCR, use the most recent data with PCR). Rationale: focus the comparison across the 14 locations rather than within-study changes.

* Restrict to clusters with at least 15 observations for serology, PCR, or TF. Rationale: since the analysis focuses on cluster-level estimates, we were concerned about estimating means or seroconversion rates with fewer observations.  The choice of 15 obs per cluster was pragmatic and based on examination of the exclusion criteria in charts (below).

This script downloads de-identified, harmonized data that have been made publicly available through the Open Science Framework as part of a larger research agenda focused on the use of serology for trachoma elimination. https://osf.io/ykjc4/ 

It creates two files: an individual level dataset `trachoma_sero_transmission_analysis_indiv.csv/.rds` and a cluster level dataset `trachoma_sero_transmission_analysis_cluster.csv/.rds`.  Note that the dataset includes some PCR and TF results that were only available at the cluster level (not individual level due to pooling) and so the cluster level dataset includes more children for PCR and TF indicators. All Pgp3 antibody data are available at the individual level.

Please refer to codebooks and metadata stored with the original datasets for additional details: https://osf.io/vj8ts/

## Preamble

```{r setup, message = FALSE}
library(here)
source(here("R/0-config.R"))
```

## Download public datasets

Download and read-in the trachoma serology harmonized datasets (individual and cluster level)

```{r read data}
#---------------------------------
# read in the harmonized,
# trachoma serology
# individual level dataset
#
# This public dataset is available
# through the Open Science Framwork
# use the OSF API through the osfr
# package to download it into the 
# local /data directory
#
# drop the "_public" suffix from
# ID variables
#---------------------------------

# download the individual level dataset (V2)
# trachoma_serology_public_data_indiv_v2.rds
# https://osf.io/3qupf
ind_osfinfo <- osf_retrieve_file("3qupf") %>%
  osf_download(path=here("data"), conflicts = "overwrite", progress = TRUE)

ind_df <- read_rds(file=ind_osfinfo$local_path) %>%
  rename(cluster_id = cluster_id_public,
         household_id = household_id_public,
         individual_id = individual_id_public) %>%
  # filter to included studies
  filter(study_id %in% c("Kongwa2013","Kongwa2018","Malawi2014","MORDOR2015","Morocco2019","PRET2013","TAITU2018","TCC-Ethiopia2017","WUHA2016")) %>%
  mutate(location_year_name = droplevels(location_year_name)) 

#---------------------------------
# read in the trachoma serology
# cluster level dataset
#
# This public dataset is available
# through the Open Science Framwork
# use the OSF API through the osfr
# package to download it into the 
# local /data directory
#
# drop the "_public" suffix from
# ID variables
#---------------------------------

# download the cluster level dataset (V2)
# trachoma_serology_public_data_cluster_v2.rds
# https://osf.io/s62p4
clu_osfinfo <- osf_retrieve_file("s62p4") %>%
  osf_download(path=here("data"), conflicts = "overwrite", progress = TRUE)

clu_df <- read_rds(file=clu_osfinfo$local_path) %>%
  rename(cluster_id = cluster_id_public) %>%
  # filter to included studies
  filter(study_id %in% c("Kongwa2013","Kongwa2018","Malawi2014","MORDOR2015","Morocco2019","PRET2013","TAITU2018","TCC-Ethiopia2017","WUHA2016")) %>%
  mutate(location_year_name = droplevels(location_year_name)) 

```


## Restrict to analysis dataset

Restrict to the most recent year of data for studies with multiple years of data. These include the WUHA trial (2016-2019, use 2019), the MORDOR trial (2015-2018, use 2018), and the Kongwa, Tanzania trial (Wilson 2019) (2012-2015, use 2013 — the most recent survey with PCR measurements).

```{r filter data to recent year}
#---------------------------------
# add flag for most recent data from each study
# NOTE: for Kongwa (Wilson 2019), 
# we are selecting 2013 because that 
# is the most recent year with PCR data. 
# data from 2014 and 2015 were also collected, 
# but using 2013 for consistency across analyses.
#---------------------------------
clu_df2 <- clu_df %>%
  mutate(most_recent = ifelse(location_year_name %in% c("Kongwa, Tanzania 2014", "Kongwa, Tanzania 2015",
                                                        "Wag Hemra, Ethiopia (WUHA) 2016", "Wag Hemra, Ethiopia (WUHA) 2017", "Wag Hemra, Ethiopia (WUHA) 2018",
                                                         "Dosso, Niger (MORDOR) 2015", "Dosso, Niger (MORDOR) 2016", "Dosso, Niger (MORDOR) 2017"),
                              0, 1))

#---------------------------------
# tabulate the number of clusters by
# study, year, and whether they 
# are included
#---------------------------------
table(clu_df2$location_year_name, clu_df2$most_recent)

clu_df3 <- clu_df2 %>%
  filter(most_recent==1) %>%
  select(-most_recent)

```

This restriction to a single year of data reduces the cluster-level dataset from `r nrow(clu_df2)` to `r nrow(clu_df3)` cluster-level summaries.


Examine influence of progressive cluster size restrictions on overall sample sizes (clusters, children).  Ideally, we'd like to include clusters with as many children as possible since the majority of this analysis relies on cluster-level summaries.  Examine restrictions along 30 (ideal) to 10 (very minimum).

```{r examine influence of cluster size restriction}
#---------------------------------
# loop over cluster restrictions
# from 5 to 30 and store the 
# resulting sample sizes
#---------------------------------

clusizes <- seq(5,30,by=1)
ss_scenarios <- foreach(clusti = clusizes, .combine=rbind) %do% {
  di <- clu_df3 %>%
    mutate(pgp3_minobs = ifelse(pgp3_n_tested >= clusti,1,0)) %>%
    filter(pgp3_minobs == 1) %>%
    mutate(ncl = 1) %>%
    summarize(n_cl=sum(ncl, na.rm=TRUE),
              n_pgp3 = sum(pgp3_n_tested, na.rm=TRUE),
              n_pcr  = sum(pcr_n_tested, na.rm=TRUE),
              n_tf   = sum(tf_n_tested, na.rm=TRUE)
              ) %>%
    mutate(minclust = clusti)
  return(di)

  
}
```

```{r plot ss scenarios clusters, fig.width=6}
#---------------------------------
# plot number of clusters by
# restriction
#---------------------------------
ggplot(data=ss_scenarios, aes(x=minclust, y = n_cl)) + 
  geom_vline(xintercept = c(15,20), color = "gray50") +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks=seq(4,30,by=2)) +
  labs(x = "minimum cluster size", y = "number of clusters included in analysis", title = "Influence of minimum cluster size on number of clusters") +
  theme_minimal()
```

```{r plot ss scenarios children, fig.width=6, fig.height=12}
#---------------------------------
# plot number of children by
# restriction
#---------------------------------
ss_scenarios2 <- ss_scenarios %>%
  select(-n_cl) %>%
  pivot_longer(cols=starts_with("n_"), values_to = "n_children", names_to = "pop")
ggplot(data=ss_scenarios2, aes(x=minclust, y = n_children)) + 
  facet_grid(pop~., scales = "free_y") +
  geom_vline(xintercept = c(15,20), color = "gray50") +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks=seq(4,30,by=2)) +
  labs(x = "minimum cluster size", y = "number of children included in analysis", title = "Influence of minimum cluster size on number of children") +
  theme_minimal()
```


```{r set minimum number of observations per cluster}
#---------------------------------
# set the minimum number of 
# observations per cluster
# based on examining the curves (above)
#---------------------------------
minobs <- 15
```

Restrict to clusters with at least `r minobs` children measured by Pgp3 antibody response.  

```{r restrict to clusters with minobs}
#---------------------------------
# Identify clusters with at least
# the min number of 
# observations for Pgp3
#---------------------------------
clu_df4 <- clu_df3 %>%
  mutate(pgp3_minobs = ifelse(pgp3_n_tested >= minobs & !is.na(pgp3_n_tested),TRUE,FALSE),
         pgp3_minobs_lab = ifelse(pgp3_n_tested >= minobs,
                              paste0("Pgp3 obs ≥",minobs),
                              paste0("Pgp3 obs <",minobs)
                              )
         ) 

```

The restricted dataset has `r nrow(clu_df4 %>% filter(pgp3_minobs))` clusters with `r format(sum(clu_df4$pgp3_n_tested[clu_df4$pgp3_minobs], na.rm=TRUE),big.mark=",")` children tested for Pgp3 IgG, `r format(sum(clu_df4$pcr_n_tested[clu_df4$pgp3_minobs], na.rm=TRUE),big.mark=",")` children tested for PCR, and `r format(sum(clu_df4$tf_n_tested, na.rm=TRUE),big.mark=",")`children tested for TF.

## Summarize the distribution of cluster sizes for each indicator, restricted to clusters with `r print(minobs)` children measured for Pgp3.

```{r plot cluster size distributions, fig.width=6, fig.height=12}
#---------------------------------
# plot number of children by
# measurement, re
#---------------------------------
plot_cl <- clu_df4 %>%
  filter(pgp3_minobs) %>%
  select(cluster_id,pgp3_n_tested, pcr_n_tested, tf_n_tested) %>%
  pivot_longer(cols=ends_with("n_tested"), values_to = "n_children", names_to = "pop")
ggplot(data=plot_cl, aes(x= n_children)) + 
  facet_grid(pop~.) +
  geom_histogram(alpha=0.3,color="black", binwidth=10) +
  scale_x_continuous(breaks=seq(0,320,by=40))+
  labs(x = "cluster size", y = "number of clusters", title = "Distribution of cluster sizes") +
  theme_minimal()
```

## Save cluster analysis dataset

Restrict to final variables and save the analysis dataset

```{r save analysis dataset}
#---------------------------------
# restrict to final variables and
# save the analysis dataset
#---------------------------------
clu_df5 <- clu_df4 %>%
  select(study_id,year,location_year_name,cluster_id,mda,
         starts_with("pcr_"),
         starts_with("tf_"),
         starts_with("ti_"),
         starts_with("pgp3_")) %>%
   mutate(location_year_name = factor(location_year_name, levels=location_year_name_order),
         location_year_name = droplevels(location_year_name))

summary(clu_df5)
str(clu_df5)

write_csv(clu_df5, file=here("data","trachoma_sero_transmission_analysis_cluster.csv"))
write_rds(clu_df5, file=here("data","trachoma_sero_transmission_analysis_cluster.rds"))
```

### Summarize N clusters and N children

The restrictions of including ≥15 children per cluster for each measurement means that there are some clusters from the total `r nrow(clu_df5)` clusters that will be excluded from different analyses. Tally the number of clusters and children available for each biomarker.

```{r tally number of clusters and children available}

#---------------------------------
# summarize marginal totals
#---------------------------------
ncl_pgp3 <- length(clu_df5$pgp3_n_tested[clu_df5$pgp3_minobs])
nch_pgp3 <- sum(clu_df5$pgp3_n_tested[clu_df5$pgp3_minobs])
ncl_pcr <- length(clu_df5$pcr_n_tested[clu_df5$pcr_n_tested>0 & clu_df5$pgp3_minobs])
nch_pcr <- sum(clu_df5$pcr_n_tested[clu_df5$pcr_n_tested>0 & clu_df5$pgp3_minobs])
ncl_tf  <- length(clu_df5$tf_n_tested[clu_df5$tf_n_tested>0 & clu_df5$pgp3_minobs])
nch_tf  <- sum(clu_df5$tf_n_tested[clu_df5$tf_n_tested>0 & clu_df5$pgp3_minobs])
ncl_ch <- data.frame(biomarker = c("Pgp3 IgG","PCR","TF"),
                     ncl = c(ncl_pgp3,ncl_pcr,ncl_tf),
                     nch = c(nch_pgp3,nch_pcr,nch_tf))
kbl(ncl_ch, 
    col.names = c("Biomarker","N clusters","N children"),
    caption = "Number of clusters with ≥15 children measured for Pgp3 per cluster and number of children in those clusters, by trachoma biomarker.",
    format.args=list(big.mark=",")) %>%
  kable_styling(bootstrap_options ="striped")

```

## Restrict individual level dataset

For the individual level dataset, restrict to child age, MDA status, clinical measures (TF/TI), PCR test results and Pgp3 antibody measurements.  Note that there are some studies for which individual-level PCR and clinical data are not available at the individual level.  Individual level data are included to estimate age-seroprevalence curves and seroconversion rates using age-seroprevalence (current status survival analysis).

```{r restrict indiv dataset}
#---------------------------------
# restrict the individual level
# dataset to the same clusters
# and study years
#---------------------------------
selected_clusters <- unique(clu_df5$cluster_id)
ind_df2 <- ind_df %>%
  left_join(clu_df5 %>% select(study_id,location_year_name,cluster_id,pgp3_minobs), by=c("study_id","location_year_name","cluster_id")) %>%
  filter(cluster_id %in% selected_clusters) %>%
  select(study_id,cluster_id,household_id,individual_id,survey,year,
         location_name,location_year_name,
         mda,
         age_months, age_years, pcr, tf, ti, 
         starts_with("pgp3_")
         ) %>%
  mutate(location_year_name = factor(location_year_name, levels=location_year_name_order),
         location_year_name = droplevels(location_year_name))

#---------------------------------
# confirm that the number of clusters
# is consistent between the two files
# strangely an NA is appearing in the 
# first value of ind_df2$cluster_id
# even though no values are missing
# subtract off 1 to account for this (!)
# confirmed there are 459, though!
#---------------------------------
length(unique(ind_df2$cluster_id[ind_df2$pgp3_minobs]))-1
nrow(clu_df5 %>% filter(pgp3_minobs))
length(unique(ind_df2$cluster_id[ind_df2$pgp3_minobs]))-1 == nrow(clu_df5 %>% filter(pgp3_minobs))

#---------------------------------
# confirm that all Pgp3 antibody
# data are consistent between the
# two files. The cluster-level dataset
# includes more observations for PCR
# and TF/TI
# 
# these numbers are limited to
# clusters with >= 15 Pgp3 measures
#---------------------------------
sum(!is.na(ind_df2$pgp3_pos[ind_df2$pgp3_minobs]))
sum(clu_df5$pgp3_n_tested[clu_df5$pgp3_minobs], na.rm=TRUE)
sum(!is.na(ind_df2$pgp3_pos[ind_df2$pgp3_minobs])) == sum(clu_df5$pgp3_n_tested[clu_df5$pgp3_minobs], na.rm=TRUE)

#---------------------------------
# list number of TF and PCR measures
# in the individual level dataset
# these Ns will be less than the 
# Ns listed in the table above
# for the cluster-level dataset
# because there are  a few studies
# that had only cluster-level data
# for PCR and TF
#---------------------------------
sum(!is.na(ind_df2$tf[ind_df2$pgp3_minobs]))
sum(!is.na(ind_df2$pcr[ind_df2$pgp3_minobs]))
```

```{r save indiv analysis dataset}
#---------------------------------
# save the individual level 
# analysis dataset
#---------------------------------
summary(ind_df2)
str(ind_df2)

write_csv(ind_df2, file=here("data","trachoma_sero_transmission_analysis_indiv.csv"))
write_rds(ind_df2, file=here("data","trachoma_sero_transmission_analysis_indiv.rds"))
```
 
## Create a summary of study-level characteristics

This study-level information is re-used in figures.

The study-level information is limited to clusters with ≥15 Pgp3 measurements.

```{r summarize by location year}
#------------------------------
# study level summary
#------------------------------
lyn_all_summary <- clu_df5 %>% 
  filter(pgp3_minobs) %>%
  group_by(study_id, year,location_year_name) %>% 
  summarise(pcr_n_tested = sum(pcr_n_tested, na.rm = TRUE),
            tf_n_tested = sum(tf_n_tested, na.rm = TRUE),
            ti_n_tested = sum(ti_n_tested, na.rm = TRUE),
            pgp3_n_tested = sum(pgp3_n_tested, na.rm = TRUE),
            pcr_prev_mean = mean(pcr_prev, na.rm = TRUE),
            pcr_prev_median = median(pcr_prev, na.rm = TRUE),
            tf_prev_mean = mean(tf_prev, na.rm = TRUE),
            tf_prev_median = median(tf_prev, na.rm = TRUE),
            ti_prev_mean = mean(ti_prev, na.rm = TRUE),
            ti_prev_median = median(ti_prev, na.rm = TRUE),
            pgp3_prev_mean = mean(pgp3_prev, na.rm = TRUE),
            pgp3_prev_median = median(pgp3_prev, na.rm = TRUE),
            pgp3_log10mean_mean = mean(pgp3_log10mean, na.rm = TRUE),
            pgp3_log10mean_median = median(pgp3_log10mean, na.rm = TRUE),
            pgp3_log10geomean_mean = mean(pgp3_log10geomean, na.rm = TRUE),
            pgp3_log10geomean_median = median(pgp3_log10geomean, na.rm = TRUE),
            .groups = "drop") %>%
  # set prev and mean IgG to NA if no children were measured
    mutate(pcr_prev_mean = ifelse(pcr_n_tested == 0, NA, pcr_prev_mean),
           pcr_prev_median = ifelse(pcr_n_tested == 0, NA, pcr_prev_median),
           tf_prev_mean = ifelse(tf_n_tested == 0, NA, tf_prev_mean),
           tf_prev_median = ifelse(tf_n_tested == 0, NA, tf_prev_median),
           ti_prev_mean = ifelse(ti_n_tested == 0, NA, ti_prev_mean),
           ti_prev_median = ifelse(ti_n_tested == 0, NA, ti_prev_median),
           pgp3_prev_mean = ifelse(pgp3_n_tested == 0, NA, pgp3_prev_mean),
           pgp3_prev_median = ifelse(pgp3_n_tested == 0, NA, pgp3_prev_median),
           pgp3_log10mean_mean = ifelse(pgp3_n_tested == 0 | is.nan(pgp3_log10mean_mean), NA, pgp3_log10mean_mean),
           pgp3_log10mean_median = ifelse(pgp3_n_tested == 0 | is.nan(pgp3_log10mean_median), NA, pgp3_log10mean_median),
           pgp3_log10geomean_mean = ifelse(pgp3_n_tested == 0 | is.nan(pgp3_log10geomean_mean), NA, pgp3_log10geomean_mean),
           pgp3_log10geomean_median = ifelse(pgp3_n_tested == 0 | is.nan(pgp3_log10geomean_median), NA, pgp3_log10geomean_median)
    )

```

```{r summarize study-level characteristics, message = FALSE}
# count total sample size and cluster size for each location-year
# number of individuals in location & year with any serology result
# limit the totals to the clusters with >= 15 children tested for Pgp3
n_df <- ind_df2 %>% 
  filter(!is.na(pgp3_pos) & pgp3_minobs) %>% 
  group_by(location_year_name, cluster_id) %>%
  mutate(nch = n()) %>%
  filter(nch>=15) %>%
  ungroup() %>%
  count(location_year_name, name = "n") 

# number of clusters in location & year with any serology result
# limit the totals to clusters with >= 15 children tested for Pgp3
m_df <- ind_df2 %>% 
  filter(pgp3_minobs) %>%
  group_by(location_year_name, cluster_id) %>% 
  summarise(pgp3_n = sum(!is.na(pgp3_pos))) %>% 
  count(location_year_name, name = "m") 

# create dataset of location-year characteristics
study_df <- ind_df2 %>% 
  distinct(location_year_name, location_name) %>% 
  left_join(lyn_all_summary, by="location_year_name") %>%
  # add label for endemicity category
  mutate(endemic_cat = case_when(
    location_name %in% c("Wag Hemra, Ethiopia (WUHA)", "Wag Hemra, Ethiopia (TAITU)", "Andabet, Ethiopia","Matameye, Niger (PRET)") ~ "High seroprevalence",
    location_year_name == "Kongwa, Tanzania 2018" ~ "Intermediate seroprevalence",
    location_name == "Kongwa, Tanzania"  ~ "High seroprevalence",
    location_name %in% c("Dera, Ethiopia", "Mchinji, Malawi", "Chikwawa, Malawi","Agdaz, Morocco") ~ "Intermediate seroprevalence",
    location_name %in% c("Alefa, Ethiopia", "Woreta town, Ethiopia", "Dosso, Niger (MORDOR)", "Boumalne Dades, Morocco", "Agdaz, Morocco") ~ "Low seroprevalence")) %>% 
  # append sample size and number of clusters to location_year_name
  left_join(n_df, by = "location_year_name") %>% 
  left_join(m_df, by = "location_year_name") %>% 
  mutate(location_year_name_nm = paste0(gsub(",", ",\n", location_year_name), "\n", "(n=", n, ", m=", m, ")")) %>% 
  # denote location / year combination that were surveillance surveys meant to be representative of study area
  mutate(location_year_name_nm = ifelse(location_year_name %in% c("Alefa, Ethiopia 2017", "Andabet, Ethiopia 2017", "Dera, Ethiopia 2017", "Woreta town, Ethiopia 2017",
                                                                  "Kongwa, Tanzania 2018", "Chikwawa, Malawi 2014", "Mchinji, Malawi 2014",
                                                                  "Boumalne Dades, Morocco 2019", "Agdaz, Morocco 2019"),
                                        paste0(location_year_name_nm, "*"), location_year_name_nm)) %>% 
  # denote cases where data collected among 1-5 only
  mutate(location_year_name_nm = ifelse(location_name %in% c("Wag Hemra, Ethiopia (TAITU)", "Dosso, Niger (MORDOR)", "Matameye, Niger (PRET)"),
                                        paste0(location_year_name_nm, " †"), location_year_name_nm)) %>% 
  # denote CC only (district-level PCR, and for 1-5yo)
  mutate(location_year_name_nm = ifelse(location_year_name %in% c("Alefa, Ethiopia 2017", "Andabet, Ethiopia 2017", "Dera, Ethiopia 2017", "Woreta town, Ethiopia 2017"),
                                        paste0(location_year_name_nm, " ‡"), location_year_name_nm)) %>% 
  mutate(location_year_name = factor(location_year_name, levels = location_year_name_order))

study_df
```

```{r save study analysis dataset}
#---------------------------------
# save the study level 
# analysis dataset
#---------------------------------
summary(study_df)
str(study_df)

write_csv(study_df, file=here("data","trachoma_sero_transmission_analysis_study.csv"))
write_rds(study_df, file=here("data","trachoma_sero_transmission_analysis_study.rds"))
```

## Session Info
```{r session info}
sessionInfo()
```