---
title: "Monitoring transmission intensity of trachoma with serology"
subtitle: "Estimate force of infection using seroconversion rate models"
author: "Contributors: Christine Tedijanto christine.tedijanto@ucsf.edu and Ben Arnold ben.arnold@ucsf.edu"
date: "Last updated: `r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
number_sections: false
---

**Objective:** Estimate force of infection using different types of seroconversion rate models (SIR, SIS, semiparametric).

The analysis generates cluster-level estimates and study-level estimates, which are in turn saved in the `output` directory as `.csv/.rds` files.  Later scripts read in these results to make summary figures.


## Setup

```{r setup, message = FALSE}
#---------------------------------
# source the configuration file
#---------------------------------
library(here)
source(here("R/0-config.R"))

# for models that allow sero-reversion 
# (SIS aka reversible catalytic models), 
# use seroreversion rate from West 2018 
# (based on 6.4% risk of seroreversion over 1 year)
west_sr <- 0.066
```


## Read data

```{r read in the data}
#---------------------------------
# read the individual level dataset
#
# filter to clusters with
# at least 15 children measured 
# for Pgp3 and non-missing obs
#---------------------------------
ind_df <- read_rds(here("data","trachoma_sero_transmission_analysis_indiv.rds")) %>%
  filter(pgp3_minobs) %>%
  filter(!is.na(pgp3_pos))

#---------------------------------
# read the cluster level dataset
#
# filter to clusters with
# at least 15 children measured 
# for Pgp3
#---------------------------------
clu_df <- read_rds(here("data","trachoma_sero_transmission_analysis_cluster.rds")) %>%
  filter(pgp3_minobs) 

```

```{r figure_prep}
save_figs <- FALSE

#---------------------------------
# set shared theme for figures
#---------------------------------
summary_stat_theme <- list(facet_wrap(.~location_year_name, ncol = 4),
                           theme_minimal(),
                           scale_fill_manual(values = location_year_name_colors),
                           scale_color_manual(values = location_year_name_colors),
                           theme(legend.position = "none"))
```


## Summarize data for each location-year in the analysis

This summary is used because for the semi-parametric estimates of seroconversion rates, these can only be estimated across all clusters in each survey (there is not enough information at the cluster level to fit an age-varying seroconversion rate model).

```{r summarize by location year}
lyn_all_summary <- clu_df %>% 
  group_by(study_id, year, location_year_name) %>% 
  summarise(mda = max(mda),
            pcr_n_tested = sum(pcr_n_tested, na.rm = TRUE),
            tf_n_tested = sum(tf_n_tested, na.rm = TRUE),
            ti_n_tested = sum(ti_n_tested, na.rm = TRUE),
            pgp3_n_tested = sum(pgp3_n_tested, na.rm = TRUE),
            pcr_prev_mean = mean(pcr_prev, na.rm = TRUE),
            pcr_prev_median = median(pcr_prev, na.rm = TRUE),
            tf_prev_mean = mean(tf_prev, na.rm = TRUE),
            tf_prev_median = median(tf_prev, na.rm = TRUE),
            ti_prev_mean = mean(ti_prev, na.rm = TRUE),
            ti_prev_median = median(ti_prev, na.rm = TRUE),
            pgp3_prev_mean = mean(pgp3_prev, na.rm = TRUE),
            pgp3_prev_median = median(pgp3_prev, na.rm = TRUE),
            pgp3_log10mean_mean = mean(pgp3_log10mean, na.rm = TRUE),
            pgp3_log10mean_median = median(pgp3_log10mean, na.rm = TRUE),
            pgp3_log10geomean_mean = mean(pgp3_log10geomean, na.rm = TRUE),
            pgp3_log10geomean_median = median(pgp3_log10geomean, na.rm = TRUE),
            .groups = "drop")

```


## Force of infection (FOI) estimates

### Simple catalytic model (SIR)

All force of infection model code borrows heavily from the following Github repositories from previous seroepidemiologic studies in the group:

* [mbita-schisto](https://github.com/ben-arnold/mbita-schisto/blob/master/R/4-mbita-schisto-by-age.Rmd, https://github.com/ben-arnold/mbita-schisto/blob/master/R/mbita-schisto-Functions.R)
* [enterics-seroepi](https://github.com/ben-arnold/enterics-seroepi/blob/master/R/SI-File7-Fig6-asembo-cross-sectional-FOI-estimation.Rmd)

```{r estimate sir models, max.width = 8.5}
## `get_sir_results`: function that returns FOI estimate from simple catalytic model -----
get_sir_results <- function(df){
  
  temp_df <- df %>% filter(!is.na(pgp3_pos))
  
  glm_fit <- tryCatch(glm(pgp3_pos~1, offset = log(age_years),
                          data = temp_df,
                          family = binomial(link = "cloglog")),
                      error = function(cond){return(cond$message)},
                      warning = function(cond){return(cond$message)})
  
  if(class(glm_fit)[1] == "character") {
    lambda <- NA
    lambda_lb <- NA
    lambda_ub <- NA
    minll <- NA
    sir_err_warn_msg <- glm_fit
  } else {
    glm_summary <- summary(glm_fit)
    log_lambda_se  <- sqrt(glm_summary$cov.unscaled)
    
    lambda <- as.numeric(exp(glm_fit$coefficients)) # use as.numeric to remove variable names
    lambda_lb <- as.numeric(exp(glm_fit$coefficients - 1.96*log_lambda_se))
    lambda_ub <- as.numeric(exp(glm_fit$coefficients + 1.96*log_lambda_se))
    minll <- -(logLik(glm_fit))
    sir_err_warn_msg <- NA
  }
  
  ret <- data.frame(sir_n = nrow(temp_df),
                    sir_foi = lambda,
                    sir_foi_lb = lambda_lb,
                    sir_foi_ub = lambda_ub,
                    sir_minll = minll,
                    sir_err_warn_msg)
  return(ret)
}

## run at each level of interest -----
clu_all_sir <- foreach(curr_lyn_clu = unique(ind_df$cluster_id), .combine = rbind) %dopar% { 
  get_sir_results(df = ind_df %>% filter(cluster_id == curr_lyn_clu)) %>% 
      mutate(cluster_id = curr_lyn_clu)
  }

clu_1to5_sir <- foreach(curr_lyn_clu = unique(ind_df$cluster_id), .combine = rbind) %dopar% {
  get_sir_results(df = ind_df %>% filter(age_years<=5, cluster_id == curr_lyn_clu)) %>% 
    mutate(cluster_id = curr_lyn_clu)
  }

clu_1to3_sir <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
  get_sir_results(df = ind_df %>% filter(age_years<=3, location_year_name == curr_lyn)) %>% 
    mutate(location_year_name = curr_lyn)
  }

lyn_all_sir <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
  get_sir_results(df = ind_df %>% filter(location_year_name == curr_lyn)) %>% 
    mutate(location_year_name = curr_lyn)
  }

lyn_all_sir <- lyn_all_sir %>%
  left_join(clu_all_sir %>%
              left_join(ind_df %>% distinct(location_year_name, cluster_id), by = "cluster_id") %>% 
              group_by(location_year_name) %>%
              summarise(sir_foi_mean = mean(sir_foi, na.rm = TRUE),
                        sir_foi_wtmean = weighted.mean(sir_foi, w = sir_n, na.rm = TRUE),
                        sir_foi_median = median(sir_foi, na.rm = TRUE),
                        .groups = "drop"),
            by = "location_year_name")
```

```{r, fig.width = 8.5, fig.height = 7}
clu_all_sir %>% 
  filter(!is.na(sir_foi)) %>% 
  left_join(ind_df %>% distinct(location_year_name, cluster_id), by = "cluster_id") %>% 
  ggplot() +
  geom_histogram(aes(x = sir_foi, fill = location_year_name),
                 color = "black", binwidth = 0.02) +
  labs(x = "cluster-level seroconversion (simple catalytic)") +
  coord_cartesian(xlim = c(0,1)) +
  summary_stat_theme
```

### Reversible catalytic model (SIS)

```{r sis model functions, max.width = 8.5}
# log-likelihood function for reversible catalytic model with fixed seroreversion rate
loglik_sis <- function(h,r,data) {
  # h    : \lambda seroconversion rate (to be estimated)
  # r    : \rho seroreversion rate (fixed)
  # data  : data frame with 1 row per age group. cols = age / n / k
  h <- rep(h,nrow(data))
  r <- rep(r,nrow(data))
  t <- data[,1]
  n <- data[,2]
  k <- data[,3]
  p <- h/(r+h)*(1-exp(-(r+h)*t))
  # negative log likelihood function (to minimize with optim)
  sum( - (k)*log(p) - (n-k)*log(1-p) )
}

## `get_sis_results`: return FOI estimates under a reversible catalytic model -----
# model with assumed seroreversion rate
# note that inference here underestimates variability since there is uncertainty in seroreversion rate
get_sis_results <- function(df, curr_sr){
  
  #print(paste0(unique(df$cluster_id), ", ", curr_sr))
  
  temp_df <- df %>% 
    filter(!is.na(pgp3_pos)) %>% 
    group_by(age_years) %>%
    summarise(pgp3_pos = sum(pgp3_pos),
              n = n(),
              .groups = "drop") %>% 
    dplyr::select(age_years, n, pgp3_pos)
  
  # maximum likelihood solution
  # with fixed sero-reversion rate
  minll <- tryCatch(optim(par = c(0.1),
                          fn = loglik_sis,
                          r = curr_sr,
                          data = temp_df,
                          hessian = TRUE,
                          method = "Brent",
                          lower = 0,
                          upper = 1),
                    error = function(cond){return(cond$message)},
                    warning = function(cond){return(cond$message)})
  
  if(class(minll)[1] == "character") {
    minll_value <- NA
    lambda_hat <- NA
    lambda_lb <- NA
    lambda_ub <- NA
    sis_err_warn_msg <- minll
  } else {
    minll_value <- minll$value
    lambda_hat <- minll$par
    # estimate SE and CI from the inverse information matrix
    I <- solve(minll$hessian)
    lambda_se <- sqrt(diag(I))
    lambda_lb <- lambda_hat-1.96*lambda_se
    lambda_ub <- lambda_hat+1.96*lambda_se
    sis_err_warn_msg <- NA
  }
  
  ret <- data.frame(seroreversion = curr_sr,
                    sis_minll = minll_value,
                    sis_foi = lambda_hat,
                    sis_foi_lb = lambda_lb,
                    sis_foi_ub = lambda_ub,
                    sis_err_warn_msg)
  return(ret)
}
```

```{r estimate sis models}
## run at each level of interest -----
clu_all_sis <- foreach(curr_lyn_clu = unique(ind_df$cluster_id), .combine = rbind) %:%
  foreach(curr_sr = c(west_sr, seq(0,0.2,0.02)), .combine = rbind) %dopar% {
    get_sis_results(df = ind_df %>% filter(cluster_id == curr_lyn_clu), curr_sr = curr_sr) %>% mutate(cluster_id = curr_lyn_clu)
    }

clu_1to5_sis <- foreach(curr_lyn_clu = unique(ind_df %>% filter(age_years<=5) %>% pull(cluster_id)), .combine = rbind) %:%
  foreach(curr_sr = c(west_sr, seq(0,0.2,0.02)), .combine = rbind) %do% {
    get_sis_results(df = ind_df %>% filter(age_years<=5, cluster_id == curr_lyn_clu), curr_sr = curr_sr) %>% mutate(cluster_id = curr_lyn_clu)
    }

lyn_all_sis <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %:%
  foreach(curr_sr = c(west_sr, seq(0,0.2,0.02)), .combine = rbind) %dopar% {
    get_sis_results(df = ind_df %>% filter(location_year_name == curr_lyn), curr_sr = curr_sr) %>% mutate(location_year_name = curr_lyn)
  } 

lyn_all_sis <- lyn_all_sis %>%
    left_join(clu_all_sis %>%
              left_join(ind_df %>% distinct(location_year_name, cluster_id), by = "cluster_id") %>% 
              group_by(location_year_name, seroreversion) %>%
              summarise(sis_foi_mean = mean(sis_foi, na.rm = TRUE),
                        sis_foi_median = median(sis_foi, na.rm = TRUE),
                        .groups = "drop"),
            by = c("location_year_name", "seroreversion"))
```

```{r, fig.width = 8.5, fig.height = 7}
clu_all_sis %>% 
  filter(seroreversion == west_sr, !is.na(sis_foi)) %>% 
  left_join(ind_df %>% distinct(location_year_name, cluster_id), by = "cluster_id") %>% 
  ggplot() +
  geom_histogram(aes(x = sis_foi, fill = location_year_name),
                 color = "black", binwidth = 0.05) +
  labs(x = "cluster-level seroconversion (reversible catalytic, sr = 0.066)") +
  coord_cartesian(xlim = c(0,3)) +
  summary_stat_theme
```

### Semiparametric spline

```{r cluster level gam}
avgFOI <- function(m, newdata, a1, a2, nreps = 10000) {

  # limit the prediction data to the two time points
  newdata <- newdata %>%
    filter(age_years %in% c(a1,a2)) %>%
    arrange(age_years)
  
  # get predicted seroprevalence, Ft, from GAM model
  gFt <- predict(m, newdata)
  Ft <- 1 - exp(-exp(gFt))  # complementary log-log link
  #Ft <- exp(gFt)/(1+exp(gFt)) # alternative logistic link (not used)
  
  # average FOI over a1 to a2 is: 
  # foi = [log(1-F(a1))-log(1-F(a2))] / (a2-a1)
  mufoi <- (log(1-Ft[1]) - log(1-Ft[2])) / (a2-a1) 

  # parametric bootstrap simulation from the
  # model coefficient estimates, assuming the model is true
  X0 <- predict(m, newdata, type = "lpmatrix")
  Vb <- vcov(m, unconditional = TRUE)
  set.seed(123)
  sims <- MASS::mvrnorm(n = nreps, mu = coef(m), Sigma = Vb)
  gFt.bs <- X0 %*% t(sims)
  Ft.bs <- 1 - exp(-exp(gFt.bs))
  #Ft.bs <- exp(gFt.bs)/(1+exp(gFt.bs)) # changed 07/19/22
  mufoi.bs <- (log(1-Ft.bs[1,]) - log(1-Ft.bs[2,])) / (a2-a1)
  
  # estimate approximate bs SE and percentile-based 95% credible interval
  mufoi.se <- sd(mufoi.bs)
  mufoi.ci <- quantile(mufoi.bs, probs = c(0.025, 0.975), na.rm = TRUE)
  ret <- data.frame(gam_foi = mufoi, gam_foi_se = mufoi.se, gam_foi_lb = mufoi.ci[1], gam_foi_ub = mufoi.ci[2])
  return(ret)
}

## `get_gam_results`: get FOI estimates from GAM (semiparametric spline) -----
get_gam_results <- function(df){
  
  temp_df <- df %>% filter(!is.na(pgp3_pos))
  
  temp_fit <- tryCatch(mgcv::gam(pgp3_pos~s(age_years, bs = "cr", k = 3),
                           data = temp_df,
                           family = binomial(link = "cloglog")),
                           #family = binomial(link = "logit")), # changed 07/19/22
                       error = function(cond){return(cond$message)},
                       warning = function(cond){return(cond$message)})

  new_df <- data.frame(age_years = c(1:9))
  
  if(class(temp_fit)[1] == "character") {
    ret <- data.frame(gam_n = nrow(temp_df),
                      gam_foi = NA,
                      gam_foi_se = NA,
                      gam_foi_lb = NA,
                      gam_foi_ub = NA,
                      gam_minll = NA,
                      gam_aic = NA,
                      gam_err_warn_msg = temp_fit)
  } else {
    ret <- avgFOI(m = temp_fit, newdata = new_df, a1 = 1, a2 = max(temp_df$age_years)) %>% 
      mutate(gam_n = nrow(temp_df),
             gam_minll = -logLik(temp_fit),
             gam_aic = AIC(temp_fit),
             gam_err_warn_msg = NA)
  }
  
  return(ret)
}

## run only at the study/district level 
## because requires a large number of obs to estimate the spline -----

lyn_all_gam <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
    get_gam_results(df = ind_df %>% filter(location_year_name == curr_lyn)) %>%
    mutate(location_year_name = curr_lyn)
    } 
```


We can also visualize the GAM curves:

```{r, fig.width = 8.5, fig.height = 8}
get_full_gam_results <- function(df) {
  eps <- 1e-07 # small value, to estimate FOI along the curve
  
  temp_df <- df %>% filter(!is.na(pgp3_pos))
  
  temp_fit <- tryCatch(mgcv::gam(pgp3_pos~s(age_years, bs = "cr", k = 3),
                           data = temp_df,
                           family = binomial(link = "logit")),
                       error = function(cond){return(cond$message)},
                       warning = function(cond){return(cond$message)})

  if(class(temp_fit)[1] == "character") {
    ret <- data.frame(age_years = NA,
                      age_foi = NA)
  } else {
    new_df <- data.frame(age_years = seq(1, max(temp_df$age_years), by = 0.1))
    new_df_eps <- new_df + eps
    
    pred_probs <- predict(temp_fit, new_df, type = "response")
    deriv <- (predict(temp_fit, new_df_eps) - predict(temp_fit, new_df)) / eps
    
    ret <- data.frame(age_years = seq(1, max(temp_df$age_years), by = 0.1),
                      age_foi = deriv * pred_probs)
  }
  return(ret)
}

## run at each level of interest -----
lyn_all_gam_full <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
  get_full_gam_results(df = ind_df %>% filter(location_year_name == curr_lyn)) %>% 
    mutate(location_year_name = curr_lyn)
  }

## visualize -----
lyn_all_gam_full %>% 
  ggplot() +
  geom_line(aes(x = age_years, y = age_foi, color = location_year_name)) +
  geom_point(data = . %>% filter(age_years %in% c(1:9)),
             aes(x = age_years, y = age_foi, fill = location_year_name),
             pch = 21, color = "black", size = 3) +
  scale_color_manual(values = location_year_name_colors) +
  scale_fill_manual(values = location_year_name_colors) +
  scale_x_continuous(breaks = c(0:9), labels = c(0:9)) +
  coord_cartesian(ylim = c(-0.1, 0.3)) +
  facet_wrap(.~location_year_name, ncol = 4) +
  labs(y = "age-specific seroconversion (semiparametric spline)", x = "age (years)") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Consolidate FOI results


```{r join cluster level estimates}
# cluster-level, all children
clu_all_summary_foi <- clu_df %>% 
  left_join(clu_all_sir, by = "cluster_id") %>% 
  left_join(clu_all_sis %>% filter(seroreversion == west_sr), by = "cluster_id") %>% 
  mutate_at(vars(sir_foi, sis_foi), ~ifelse(pgp3_prev == 0, 0, .)) %>% 
  mutate(sir_foi = ifelse(!is.na(sir_err_warn_msg), NA, sir_foi)) %>% 
  mutate(sis_foi = ifelse(!is.na(sis_err_warn_msg), NA, sis_foi)) 

# cluster-level, 1-5yo
clu_1to5_summary_foi <- clu_df %>% 
  left_join(clu_1to5_sir, by = "cluster_id") %>% 
  left_join(clu_1to5_sis %>% filter(seroreversion == west_sr), by = "cluster_id") %>% 
  mutate_at(vars(sir_foi, sis_foi), ~ifelse(pgp3_prev == 0, 0, .)) %>% 
  mutate(sir_foi = ifelse(!is.na(sir_err_warn_msg), NA, sir_foi)) %>% 
  mutate(sis_foi = ifelse(!is.na(sis_err_warn_msg), NA, sis_foi))

# cluster-level, 1-9yo
clu_1to9_summary_foi <- clu_all_summary_foi %>% 
  filter(study_id %ni% c("PRET2013", "MORDOR2015")) %>% # remove location + year combos without full 1-9yo
  mutate_at(vars(starts_with("pgp3"), starts_with("ct694"), starts_with("sir"), starts_with("sis"), starts_with("gam")),
            ~ifelse(study_id == "WUHA2016" & year %in% c(2017, 2018), NA, .)) # WUHA month 12+24 serology are 1-5yo only

```

```{r join study level estimates}

# location-year-level, all
# add number of clusters where models converged for SIR and SIS models
lyn_all_summary_foi <- lyn_all_summary %>% 
  left_join(lyn_all_sir, by = "location_year_name") %>% 
  left_join(lyn_all_sis %>% filter(seroreversion == west_sr), by = "location_year_name") %>% 
  left_join(lyn_all_gam, by = "location_year_name") %>% 
  left_join(clu_all_summary_foi %>% filter(is.na(sir_err_warn_msg)) %>% count(location_year_name, name = "sir_converge_n"),
            by = c("location_year_name")) %>% 
  left_join(clu_all_summary_foi %>% filter(is.na(sis_err_warn_msg)) %>% count(location_year_name, seroreversion, name = "sis_converge_n"),
            by = c("location_year_name", "seroreversion")) 
```

## Store estimates

```{r save estimates}
# location-year-level, all
write_csv(lyn_all_summary_foi, file = here("output", "lyn_all_summary.csv"))
write_rds(lyn_all_summary_foi, file = here("output", "lyn_all_summary.rds"))
write_csv(lyn_all_sis, file = here("output", "lyn_all_sis.csv"))
write_rds(lyn_all_sis, file = here("output", "lyn_all_sis.rds"))

# cluster-level, 1-5 only
write_csv(clu_1to5_summary_foi, file = here("output", "clu_1to5_summary.csv"))
write_rds(clu_1to5_summary_foi, file = here("output", "clu_1to5_summary.rds"))
write_csv(clu_1to5_sis, file = here("output", "clu_1to5_sis.csv"))
write_rds(clu_1to5_sis, file = here("output", "clu_1to5_sis.rds"))

# cluster-level, 1-9 only
write_csv(clu_1to9_summary_foi, file = here("output", "clu_1to9_summary.csv"))
write_rds(clu_1to9_summary_foi, file = here("output", "clu_1to9_summary.rds"))

# cluster-level, all
write_csv(clu_all_summary_foi, file = here("output", "clu_all_summary.csv"))
write_rds(clu_all_summary_foi, file = here("output", "clu_all_summary.rds"))
write_csv(clu_all_sis, file = here("output", "clu_all_sis.csv"))
write_rds(clu_all_sis, file = here("output", "clu_all_sis.rds"))
```

## Session Info
```{r session info}
sessionInfo()
```