---
title: "Monitoring transmission intensity of trachoma with serology"
subtitle: "Summarize study means"
author: "Contributors: Christine Tedijanto christine.tedijanto@ucsf.edu and Ben Arnold ben.arnold@ucsf.edu"
date: "Last updated: `r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
number_sections: false
---

**Objective:** Plot distributions of study cluster means for seroprevalence, PCR prevalence, and TF prevalence.


## Setup

```{r setup, message = FALSE}
#------------------------------
# Load project config file
#------------------------------
library(here)
source(here("R/0-config.R"))
```

```{r figure_prep}

#------------------------------
# global variables
#------------------------------
# set shared theme for figures
summary_stat_theme <- list(facet_wrap(.~location_year_name, ncol = 4),
                           theme_minimal(),
                           scale_fill_manual(values = location_year_name_colors),
                           scale_color_manual(values = location_year_name_colors),
                           theme(legend.position = "none"))

# minimum number of measurements per cluster
min_clu_n <- 15

# save figures to output?
save_figs <- TRUE
```

## Read data and stored estimates

```{r read data}
#------------------------------
# load analysis datasets
#------------------------------
# read individual-level dataset
# filter to clusters with ≥ 15 pgp3 measurements
ind_df <- read_rds(here("data", "trachoma_sero_transmission_analysis_indiv.rds")) %>%
  filter(pgp3_minobs)

# read cluster-level dataset
# filter to clusters with ≥ 15 pgp3 measurements
clu_df <- read_rds(here("data","trachoma_sero_transmission_analysis_cluster.rds")) %>%
  filter(pgp3_minobs) 

# read study-level dataset
study_df <- read_rds(here("data","trachoma_sero_transmission_analysis_study.rds"))

#------------------------------
# read in FOI estimates
# created by 2-estimate-foi.Rmd
#------------------------------
clu_all_summary_foi <- read_rds(here("output","clu_all_summary.rds"))
lyn_all_summary_foi <- read_rds(here("output","lyn_all_summary.rds"))

# join study-level information to the study-level FOI estimates
lyn_all_summary_foi <- lyn_all_summary_foi %>%
  left_join(study_df %>% select(location_year_name, endemic_cat, n, m, location_year_name_nm), by=c("location_year_name"))
```


## Distributions by location + year

### Fig S2. Age distribution by location + year

```{r, fig.width = 8.5, fig.height = 6}
age_hist <- ind_df %>%
  # limit to observations with Pgp3 measurements
  filter(!is.na(pgp3_bin)) %>% 
  ggplot() +
  geom_histogram(aes(x = age_years, fill = location_year_name),
           color = "black", binwidth = 1, alpha = 0.8) +
  scale_x_continuous(limits = c(0,10), breaks = c(1:9)) +
  labs(x = "Age (years)", y = "Number of children") +
  summary_stat_theme +
  theme(strip.text = element_text(size = 8))

age_hist

if(save_figs){
  ggsave(filename = here::here("output", "figS2_age_dist.png"),
         plot = age_hist,
         width = 8.5, height = 6.5)
}
```

### pgp3 response distribution by location + year

```{r, fig.width = 8.5, fig.height = 7}
pgp3_dist <- ind_df %>%
  filter(!is.na(pgp3_nonneg)) %>%
  ggplot() +
  geom_histogram(aes(x = pgp3_nonneg, fill = location_year_name),
           color = "black", binwidth = 0.15) +
  geom_vline(data = . %>% distinct(location_year_name, pgp3_cutoff),
             aes(xintercept = pgp3_cutoff)) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x)(c(1, 1e5)),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  labs(x = "Pgp3 IgG response (MFI-bg)",
       y = "Number of children") +
  summary_stat_theme +
  facet_wrap(.~location_year_name, ncol = 4, scales = "free_y")

pgp3_dist
```


### Figure S1. Cluster-level sample sizes

```{r, fig.width = 8.5, fig.height = 6.5}
clu_n <- clu_df %>%
  ggplot() +
  geom_histogram(aes(x = pgp3_n_tested, fill = location_year_name),
                 color = "black", binwidth = 5, alpha = 0.8) +
  scale_x_continuous(breaks=seq(0,300,by=50)) + 
  labs(x = "Number of children assessed for serology outcomes",
       y = "Number of clusters") +
  summary_stat_theme +
  theme(strip.text = element_text(size = 8))

clu_n

if(save_figs){
  ggsave(filename = here::here("output", "figS1_clu_sample_sizes.png"),
         plot = clu_n,
         width = 8.5, height = 6.5)
}
```

### Cluster-level PCR

```{r, fig.width = 8.5, fig.height = 4}
clu_df %>%
  filter(!is.na(pcr_prev)) %>%
  ggplot() +
  geom_histogram(aes(x = pcr_prev, fill = location_year_name),
                 color = "black", binwidth = 2) +
  geom_vline(data = study_df %>% filter(!is.na(pcr_prev_mean)), aes(xintercept = pcr_prev_mean)) +
  geom_vline(data = study_df %>% filter(!is.na(pcr_prev_median)), aes(xintercept = pcr_prev_median), lty = "dashed") +
  labs(x = "cluster-level pcr prevalence (%)") +
  summary_stat_theme
```

### Cluster-level TF

```{r, fig.width = 8.5, fig.height = 5}
clu_df %>% 
  filter(!is.na(tf_prev)) %>% 
  ggplot() +
  geom_histogram(aes(x = tf_prev, fill = location_year_name),
                 color = "black", binwidth = 2) +
  geom_vline(data = study_df %>% filter(!is.na(tf_prev_mean)), aes(xintercept = tf_prev_mean)) +
  geom_vline(data = study_df %>% filter(!is.na(tf_prev_median)), aes(xintercept = tf_prev_median), lty = "dashed") +
  geom_vline(xintercept = 5, lwd = 0.75, color = "red") +
  labs(x = "cluster-level TF prevalence (%)") +
  summary_stat_theme +
  facet_wrap(.~location_year_name, ncol = 4, scales = "free_y")
```

### Cluster-level TI

```{r, fig.width = 8.5, fig.height = 5}
clu_df %>% 
  filter(!is.na(ti_prev)) %>% 
  ggplot() +
  geom_histogram(aes(x = ti_prev, fill = location_year_name),
                 color = "black", binwidth = 3) +
  geom_vline(data = study_df %>% filter(!is.na(ti_prev_mean)), aes(xintercept = ti_prev_mean)) +
  geom_vline(data = study_df %>% filter(!is.na(ti_prev_median)), aes(xintercept = ti_prev_median), lty = "dashed") +
  labs(x = "cluster-level TI prevalence (%)") +
  summary_stat_theme +
  facet_wrap(.~location_year_name, ncol = 4, scales = "free_y")
```

### Cluster-level seroprevalence

```{r, fig.width = 8.5, fig.height = 7}
clu_df %>% 
  filter(!is.na(pgp3_prev)) %>% 
  ggplot() +
  geom_histogram(aes(x = pgp3_prev, fill = location_year_name),
                 color = "black", binwidth = 2) +
  geom_vline(data = study_df %>% filter(!is.na(pgp3_prev_mean)), aes(xintercept = pgp3_prev_mean)) +
  geom_vline(data = study_df %>% filter(!is.na(pgp3_prev_median)), aes(xintercept = pgp3_prev_median), lty = "dashed") +
  labs(x = "cluster-level pgp3 seroprevalence (%)") +
  summary_stat_theme +
  facet_wrap(.~location_year_name, ncol = 4, scales = "free_y")
```

### Cluster-level mean antibody levels

```{r, fig.width = 8.5, fig.height = 7}
clu_df %>%
  filter(!is.na(pgp3_log10mean)) %>% 
  ggplot() +
  geom_histogram(aes(x = pgp3_log10mean, fill = location_year_name),
                 color = "black", binwidth = 0.1) +
  geom_vline(data = study_df %>% filter(!is.na(pgp3_log10mean_mean)), aes(xintercept = pgp3_log10mean_mean)) +
  geom_vline(data = study_df %>% filter(!is.na(pgp3_log10mean_median)), aes(xintercept = pgp3_log10mean_median), lty = "dashed") +
  labs(x = "cluster-level anti-pgp3 antibody response (log10 MFI-bg)") +
  summary_stat_theme
```


## Figure 1

For each study, summarize the distribution of cluster-level Pgp3 seroprevalence, PCR infection prevalence, and TF prevalence.  Plot cluster level values along with box plots (25th, 50th, 75th percentiles) of the distributions.

```{r}
cc_pcr_manual <- data.frame(location_year_name = c("Andabet, Ethiopia 2017",
                                                   "Dera, Ethiopia 2017",
                                                   "Woreta town, Ethiopia 2017",
                                                   "Alefa, Ethiopia 2017"),
                            prev = c(11.3, 0, 0, 0),
                            stringsAsFactors = FALSE)

## create age-seroprevalence curve visualization -----
# 3 panels (from L to R): seroprevalence, pcr, TF
get_age_sero_curve_fig <- function(incl_legend){
  
  lyn_recent <- ind_df %>%
    left_join(study_df %>% distinct(location_year_name), by = c("location_year_name")) %>% 
    pull(location_year_name) %>%
    unique()
  
  # create shared theme
  temp_theme <- list(scale_color_manual(values = location_year_name_colors),
                     scale_fill_manual(values = location_year_name_colors),
                     guides(color = FALSE, fill = FALSE),
                     theme_minimal(),
                     theme(legend.position = "none"))
  
  # if(curr_endemic_cat == "hyperendemic (median TF >20%)"){temp_cex = 3}
  # else if(curr_endemic_cat == "mesoendemic (median TF 5-20%)"){temp_cex = 2}
  # else if(curr_endemic_cat == "hypoendemic (median TF <5%)"){temp_cex = 1.3}
  
  temp_cex <- 2
  
  # temp_lyn_order <- lyn_all_summary_foi %>% 
  #   filter(location_year_name %in% lyn_recent) %>% 
  #   #arrange(desc(sir_foi)) %>% 
  #   # sort studies by median seroprevalence
  #   # two studies have basically identical median seroprevalence 
  #   # in each of the highest transmission (TAITU, Kongwa 2013), 
  #   # and intermediate transmission (Dera Ethiopia, Agdaz Morocco) settings
  #   # but visually there is more variance in TAITU  and Dera so order them higher
  #   mutate(higher_var = ifelse(location_year_name %in% c("Wag Hemra, Ethiopia (WUHA) 2019","Andabet, Ethiopia 2017","Wag Hemra, Ethiopia (TAITU) 2018"),1,0),
  #          lower_var = ifelse(location_year_name %in% c("Agdaz, Morocco 2019"),1,0)
  #          ) %>%
  #   arrange(endemic_cat,desc(higher_var),lower_var,
  #           desc(pgp3_prev_median), desc(pcr_prev_median), desc(tf_prev_median)) %>% 
  #   pull(location_year_name)
  
  # order studies consistently with their order in Figure 2.
  # these are ordered broadly by Pgp3 Seroconversion rate
temp_lyn_order <- c(
  "Wag Hemra, Ethiopia (WUHA) 2019",
  "Wag Hemra, Ethiopia (TAITU) 2018",
  "Andabet, Ethiopia 2017",
  "Matameye, Niger (PRET) 2013",
  "Kongwa, Tanzania 2013",
  "Kongwa, Tanzania 2018",
  "Chikwawa, Malawi 2014",
  "Mchinji, Malawi 2014",
  "Dera, Ethiopia 2017",
  "Agdaz, Morocco 2019",
  "Woreta town, Ethiopia 2017",
  "Alefa, Ethiopia 2017",
  "Boumalne Dades, Morocco 2019",
  "Dosso, Niger (MORDOR) 2018"
)
  
  # 1) SP figures
  sp_fig <- clu_all_summary_foi %>%
    filter(location_year_name %in% lyn_recent) %>%
    filter(pgp3_n_tested >= min_clu_n) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>% 
    ggplot() +
    # geom_jitter(aes(x = location_year_name, y = pgp3_prev, color = location_year_name),
    #             width=0.1,alpha = 0.8, size = 1.2) +
    geom_point(aes(x = location_year_name, y = pgp3_prev, color = location_year_name),
               shape = "|", alpha = 0.8, size = 4) +
    geom_boxplot(aes(x = location_year_name, y = pgp3_prev),
                 color = "black", width = 0.3, outlier.shape = NA, fill = NA, lwd = 0.8) +
    scale_x_discrete(labels = ind_df %>%
                       left_join(lyn_all_summary_foi %>% dplyr::select(location_year_name, location_year_name_nm), by = "location_year_name") %>% 
                       distinct(location_year_name, location_year_name_nm) %>%
                       tibble::deframe(),
                     limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 100, by = 20), limits = c(0,100)) +
    labs(x = NULL, y = "Seroprevalence (%)") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(size = 12),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_text(color = rev(location_year_name_colors[temp_lyn_order]),
                                     size = 12, hjust = 0.5))
  
  # 2) FOI figures
  foi_fig <- clu_all_summary_foi %>%
    filter(location_year_name %in% lyn_recent) %>%
    filter(pgp3_n_tested >= min_clu_n) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>% 
    ggplot() +
    geom_point(aes(x = location_year_name, y = sir_foi, color = location_year_name),
               shape = "|", alpha = 0.8, size = 4) +
    geom_boxplot(aes(x = location_year_name, y = sir_foi),
                 color = "black", width = 0.3, outlier.shape = NA, fill = NA, lwd = 0.8) +
    scale_x_discrete(labels = ind_df %>%
                       left_join(lyn_all_summary_foi %>% dplyr::select(location_year_name, location_year_name_nm), by = "location_year_name") %>% 
                       distinct(location_year_name, location_year_name_nm) %>%
                       tibble::deframe(),
                     limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0,1)) +
    labs(x = NULL, y = "Seroconversion rate per person-year") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(size = 12),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_blank())
  
  # 3) PCR figures
  pcr_fig <- clu_all_summary_foi %>%
    filter(location_year_name %in% lyn_recent) %>% 
    filter(pcr_n_tested >= min_clu_n) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>%  
    ggplot() +
    geom_point(aes(x = location_year_name, y = pcr_prev, color = location_year_name),
               shape = "|", alpha = 0.8, size = 4) +
    # geom_jitter(aes(x = location_year_name, y = pcr_prev, color = location_year_name),
    #             width=0.1,alpha = 0.8, size = 1.2) +
    geom_boxplot(aes(x = location_year_name, y = pcr_prev),
                 color = "black", width = 0.3, outlier.shape = NA, fill = NA, lwd = 0.8) +
    # add manual points for Carter Center
    geom_point(data = cc_pcr_manual %>% filter(location_year_name %in% temp_lyn_order),
               aes(x = location_year_name, y = prev, fill = location_year_name),
               alpha = 0.7, pch = 21, color = "black", size = 4) +
    scale_x_discrete(limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 100, by = 20), limits = c(0,100)) +
    labs(x = NULL, y = "PCR prevalence (%)") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(size = 12),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_blank())
  
  # 4) TF figures
  tf_fig <- clu_all_summary_foi %>%
    filter(tf_n_tested >= min_clu_n | study_id == "TCC-Ethiopia2017") %>% 
    filter(location_year_name %in% lyn_recent) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>%  
    ggplot() +
    geom_point(aes(x = location_year_name, y = tf_prev, color = location_year_name),
               shape = "|", alpha = 0.7, size = 4) +
    # geom_jitter(aes(x = location_year_name, y = tf_prev, color = location_year_name),
    #             width=0.1,alpha = 0.8, size = 1.2) +
    geom_boxplot(aes(x = location_year_name, y = tf_prev),
                 color = "black", width = 0.3, outlier.shape = NA, fill = NA, lwd = 0.8) +
    scale_x_discrete(limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 100, by = 20), limits = c(0,100)) +
    labs(x = NULL, y = "TF prevalence (%)") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(size = 12),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_blank())
  
  ret <- plot_grid(sp_fig,
                   #foi_fig,
                   pcr_fig,
                   tf_fig,
                   nrow = 1,
                   rel_widths = c(1, 0.6, 0.6, 0.6))
  
  return(ret)
}
```

```{r, fig.width = 8.5, fig.height = 8.5, message = FALSE, warning = FALSE}
age_sero_combined_fig <- get_age_sero_curve_fig(incl_legend = FALSE)

age_sero_combined_fig

if(save_figs){
  ggsave(filename = here::here("output", "fig1_trach_summary.png"),
         plot = age_sero_combined_fig,
         width = 11, height = 12, units = "in")
}
```

## For text

```{r}
clu_df_elig <- clu_df %>% filter(pgp3_n_tested>=min_clu_n)
```

* Mininum number of serological measurements per cluster: `r min_clu_n`
* Total number of clusters: `r nrow(clu_df_elig)`
* Median children per cluster: `r median(clu_df_elig$pgp3_n_tested)`, IQR: `r quantile(clu_df_elig$pgp3_n_tested, probs = c(0.25))`, `r quantile(clu_df_elig$pgp3_n_tested, probs = c(0.75))`
* Total serological measurements: `r sum(clu_df_elig$pgp3_n_tested)`

* Seroprevalence median: `r round(median(study_df$pgp3_prev_median, na.rm=TRUE),0)`, range: `r round(min(study_df$pgp3_prev_median, na.rm=TRUE),0)`, `r round(max(study_df$pgp3_prev_median, na.rm=TRUE),0)`
* PCR prevalence median: `r round(median(study_df$pcr_prev_median, na.rm = TRUE),0)`, range: `r round(min(study_df$pcr_prev_median, na.rm = TRUE),0)`, `r round(max(study_df$pcr_prev_median, na.rm = TRUE),0)`
* TF prevalence median: `r round(median(study_df$tf_prev_median, na.rm = TRUE),0)`, range: `r round(min(study_df$tf_prev_median, na.rm = TRUE),0)`, `r round(max(study_df$tf_prev_median, na.rm = TRUE),0)`

Among clusters with MDA, the median PCR prevalence was `r round(median(clu_df_elig %>% filter(mda == 1) %>% pull(pcr_prev), na.rm = TRUE), 0)`.
Among clusters without recent MDA, the median PCR prevalence was `r round(median(clu_df_elig %>% filter(mda == 0) %>% pull(pcr_prev), na.rm = TRUE), 0)`.

## Session Info

```{r session info}
sessionInfo()
```
