---
title: "Monitoring transmission intensity of trachoma with serology"
subtitle: "Estimate age-seroprevalence curves"
author: "Christine Tedijanto christine.tedijanto@ucsf.edu"
date: "Last updated: `r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
number_sections: false
---

**Objective:** To create figures summarizing serological summmary measures including age-seroprevalence curves, summary scatterplots, and comparisons at different cutoffs.

## Setup


```{r setup, message = FALSE}
#------------------------------
# Load project config file
#------------------------------
library(here)
source(here("R/0-config.R"))

#------------------------------
# global variables
#------------------------------

# minimum number of measurements per cluster
# see 1-trachoma-sero-transmission-make-datasets.Rmd
# for examination of minimum cluster size. 15 chosen
min_clu_n <- 15 

# save figures?
save_figs <- TRUE 
```

## Load data

Prepare datasets and variables that are used throughout analysis. 

```{r prep_data}
#------------------------------
# load analysis datasets
#------------------------------

# read study-level dataset
study_df <- read_rds(here("data","trachoma_sero_transmission_analysis_study.rds"))

# read individual-level dataset
# filter to clusters with >= 15 Pgp3 measurements
# and non-missing Pgp3 observations
ind_df <- read_rds(here("data", "trachoma_sero_transmission_analysis_indiv.rds")) %>%
  left_join(study_df %>% select(location_year_name, endemic_cat), by ="location_year_name") %>%
  filter(pgp3_minobs & !is.na(pgp3_bin))

#------------------------------
# load cluster-level estimates
# these estimates were generated
# in the script 2-estimate-foi.Rmd
#
# join on the endemicity category info
#------------------------------
clu_all <- read_rds(here("output", "clu_all_summary.rds")) %>% 
  left_join(study_df %>% select(location_year_name, endemic_cat), by = "location_year_name")

clu_1to5 <- read_rds(here("output", "clu_1to5_summary.rds")) %>% 
  left_join(study_df %>% select(location_year_name, endemic_cat), by = "location_year_name")

#------------------------------
# load study-level estimates
# these estimates were generated
# in the script 2-estimate-foi.Rmd
#
# join on the endemicity category info
#------------------------------
lyn_all  <- read_rds(here("output", "lyn_all_summary.rds")) %>% 
  rowwise() %>% 
  mutate(lyn_short = paste(str_split(location_year_name, ",")[[1]][1], year)) %>% 
  ungroup() %>%
  left_join(study_df %>% select(location_year_name, endemic_cat), by = "location_year_name")

```

```{r figure_prep}
# set shared theme for figures
shared_theme <- list(facet_wrap(.~location_year_name, ncol = 4),
                     theme_minimal(),
                     scale_fill_manual(values = location_year_name_colors),
                     scale_color_manual(values = location_year_name_colors),
                     theme(legend.position = "none"))
```

## Age-seroprevalence curves

Carter Center cluster-level PCR data is not available (data only processed for district-level estimates). Instead, we use mean and 95% CI values from [Nash 2020](https://doi.org/10.4269/ajtmh.20-0777). Note that PCR was calculated among children aged 1-5 years, and the estimation procedures did not allow for calculation of CIs around this point estimate.

```{r cc_data_manual}
cc_pcr_manual <- data.frame(location_year_name = c("Andabet, Ethiopia 2017",
                                                   "Dera, Ethiopia 2017",
                                                   "Woreta town, Ethiopia 2017",
                                                   "Alefa, Ethiopia 2017"),
                            prev = c(11.3, 0, 0, 0),
                            stringsAsFactors = FALSE)
```

**Estimate empirical age-seroprevalence values for each location and year.**

```{r}
age_sero_results <- ind_df %>%
  group_by(study_id, location_name, year, location_year_name, endemic_cat, age_years) %>%
  summarise(pgp3_n_pos = sum(pgp3_bin == 1, na.rm = TRUE),
            pgp3_n_tested = sum(!is.na(pgp3_bin), na.rm = TRUE),
            pgp3_mean = mean(pgp3_cont, na.rm = TRUE),
            pgp3_log10mean = mean(pgp3_log10, na.rm = TRUE),
            .groups = "drop") %>% 
  mutate(pgp3_prev = pgp3_n_pos / pgp3_n_tested * 100)
```

**Quick diagnostic to hypothesize is FOI is age-dependent: plot -ln(seroprevalence) against age:**

We expect a linear relationship if FOI is constant (see [Vynnycky and White text](https://anintroductiontoinfectiousdiseasemodelling.com/)).

```{r, fig.width = 8.5}
age_sero_results %>%
  mutate(pgp3_prev_negln = ifelse(pgp3_prev == 0, 0, -log(pgp3_prev/100))) %>% 
  ggplot(aes(x = age_years, y = pgp3_prev_negln)) +
  geom_smooth(formula = y~x, color = "black",
              lwd = 0.5, alpha = 0.5, method = "lm") +
  geom_point(aes(fill = location_year_name), color = "black",
             pch = 21, size = 1.5) +
  scale_x_continuous(limits = c(0,10), breaks = c(0:10)) +
  labs(y = "-ln(anti-pgp3 seroprevalence)", x = "age (years)") +
  shared_theme +
  facet_wrap(.~location_year_name, ncol = 4, scales = "free_y")
```

### Figure 1. Age-seroprevalence curves, most recent

From Ben's notes:

* Helpful blog post on random effects and splines: https://fromthebottomoftheheap.net/2021/02/02/random-effects-in-gams/
* `mgcv::gam` and `lme4::glmer` produced fairly similar curves in a mini-test; `gam` was less wiggly

```{r}
## Example code for glmer
# library(splines)
# test_glmer <- lme4::glmer(sero_bin ~ bs(age_years, degree = 3) + (1|cluster_id),
#                           data = combined_recent_df %>% filter(location_year_name == "Kongwa, Tanzania 2015"),
#                           family = binomial(link = "logit"))
# 
# test_glmer_pred <- predict(test_glmer, newdata = new_df, re.form = ~0, se = FALSE, type = "response")
```

```{r}
## `fit_age_sero_curve` function -----
# fit age seroprevalence curves using cubic splines in `gam`
# return fitted splines and 95% CIs
age_sero_curve_results <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
  
  temp_df <- ind_df %>%
    mutate(cluster_id = as.factor(cluster_id),
           dummy = 1) %>%
    filter(location_year_name == curr_lyn)
  
  age_grid <- seq(1, max(temp_df$age_years), by = 0.1)
  
  mod <- mgcv::gam(pgp3_bin ~ s(age_years, k = 4, bs = "cr"), #+
                     #s(cluster_id, bs = "re", by = dummy),
                   data = temp_df,
                   family = binomial(link = "logit"),
                   #family = gaussian(),
                   method = "REML")
  
  k_check <- mgcv::k.check(mod)
  
  new_df <- data.frame(age_years = age_grid,
                       cluster_id = temp_df$cluster_id[1],
                       dummy = 0)
  
  preds <- predict(mod, newdata = new_df, se = TRUE)
  
  ret <- data.frame(location_year_name = curr_lyn,
                    study_id = unique(temp_df$study_id),
                    location_name = unique(temp_df$location_name),
                    year = unique(temp_df$year),
                    age_years = age_grid,
                    endemic_cat = unique(temp_df$endemic_cat),
                    preds = expit(preds$fit) * 100,
                    preds_lwrci = expit(preds$fit + qnorm(p = 0.025)*preds$se.fit) * 100,
                    preds_uprci = expit(preds$fit + qnorm(p = 0.975)*preds$se.fit) * 100,
                    k_index = k_check['s(age_years)','k-index'],
                    k_index_pval = k_check['s(age_years)','p-value'])
                    ## for gaussian fit
                    # preds = preds$fit,
                    # preds_lwrci = preds$fit + qnorm(p = 0.025)*preds$se.fit,
                    # preds_uprci = preds$fit + qnorm(p = 0.975)*preds$se.fit)
  
  return(ret)
  
}
```

```{r}
# age_sero_trendtest <- foreach(curr_lyn = unique(ind_df$location_year_name), .combine = rbind) %dopar% {
#   temp_df <- age_sero_results %>% 
#     filter(location_year_name == curr_lyn) %>% 
#     arrange(age_years)
#   
#   trend_test <- trend::mk.test(x = c(temp_df %>% pull(pgp3_prev)),
#                                alternative = "two.sided")
#   
#   data.frame(location_year_name = curr_lyn,
#              mk_test_p = trend_test$p.value,
#              mk_test_reject = ifelse(trend_test$p.value<=0.05, "p<=0.05","p>0.05"))
# }

# create shared theme
temp_theme <- list(scale_color_manual(values = location_year_name_colors),
                   scale_fill_manual(values = location_year_name_colors),
                   theme_minimal(),
                   theme(legend.position = "none"))

get_age_sero_curve <- function(pcr_min, pcr_max){
  
  temp_df <- age_sero_curve_results %>%
    left_join(cc_pcr_manual, by = "location_year_name") %>% 
    left_join(lyn_all %>% dplyr::select(location_year_name, pcr_prev_median),
              by = "location_year_name") 
    #left_join(age_sero_trendtest, by = "location_year_name")
    #mutate(pcr_prev_median = coalesce(pcr_prev_median, prev)) %>% 

  temp_df %>% 
    filter(age_years>0) %>% 
      ggplot(aes(x = age_years, y = preds, color = location_year_name)) +
      geom_line(lwd = 0.8) +
      #geom_line(lwd = 0.8, aes(lty = as.factor(mk_test_reject))) +
      # geom_point(data = age_sero_results  %>%
      #                   filter(age_years>0) %>%
      #                   filter(location_year_name %in% unique(temp_df$location_year_name)),
      #            aes(x = age_years, y = pgp3_prev,
      #                fill = location_year_name),
      #                #size = pgp3_n_tested),
      #            alpha = 0.6, size = 1) +
      # scale_linetype_manual(values = c("p>0.05" = "dashed", "p<=0.05" = "solid"),
      #                       name = "Mann-Kendall trend test") +
      #scale_radius(range = c(1,7), breaks = seq(0 , 1000, by = 250), limits = c(1, 1000)) +
      scale_x_continuous(breaks = seq(0, 9, by = 1)) +
      scale_y_continuous(breaks = seq(0, 60, by = 10)) +
      labs(x = "Age (years)", y = "Seroprevalence (%)",
          size = "Sample size") +
      guides(color = "none", fill = "none") +
      coord_cartesian(ylim = c(0,60)) +
      temp_theme +
      theme(legend.position = c(0.2, 0.9), legend.key.height = unit(0.5, "cm"),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 9)) +
      guides(linetype = guide_legend(ncol = 1))
}

age_sero_fig <- get_age_sero_curve(pcr_min = -Inf, pcr_max = Inf)

# temp_lyn_order <- lyn_all %>% 
#     #arrange(desc(sir_foi)) %>%
#     mutate(sir_foi_median = ifelse(sir_foi_median<1e-10, 0, sir_foi_median)) %>% 
#     arrange(desc(sir_foi_median), desc(pcr_prev_median), as.character(location_year_name)) %>% 
#     pull(location_year_name)

# order studies consistently with their order in Figure 1. 
# this is by their seroconversion rates
temp_lyn_order <- c(
  "Wag Hemra, Ethiopia (WUHA) 2019",
  "Wag Hemra, Ethiopia (TAITU) 2018",
  "Andabet, Ethiopia 2017",
  "Matameye, Niger (PRET) 2013",
  "Kongwa, Tanzania 2013",
  "Kongwa, Tanzania 2018",
  "Chikwawa, Malawi 2014",
  "Mchinji, Malawi 2014",
  "Dera, Ethiopia 2017",
  "Agdaz, Morocco 2019",
  "Woreta town, Ethiopia 2017",
  "Alefa, Ethiopia 2017",
  "Boumalne Dades, Morocco 2019",
  "Dosso, Niger (MORDOR) 2018"
)

## foi simple -----
sir_foi_fig <- clu_all %>%
    # restrict to clusters with at least minimum cluster size
    filter(pgp3_n_tested >= min_clu_n) %>% 
    mutate(sir_foi = sir_foi*100) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>% 
    ggplot() +
    geom_point(aes(x = location_year_name, y = sir_foi, color = location_year_name),
               shape = "|", alpha = 0.8, size = 1.5) +
    geom_boxplot(aes(x = location_year_name, y = sir_foi),
                 color = "black", width = 0.5, outlier.shape = NA, fill = NA, lwd = 0.4) +
    scale_x_discrete(labels = data.frame(location_year_name = temp_lyn_order) %>%
                       left_join(location_year_name_short, by = "location_year_name") %>% 
                       pull(lyn_short) %>% rev(), 
                     limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 70, by = 10), limits = c(0,70)) + # note: this masks a single cluster in Wag Hemra (TAITU) that is close to 70 per 100 child years.
    labs(x = NULL, y = "Seroconversion rate per 100 PY") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 7),
          axis.title.x = element_text(size = 8),
          axis.text.y = element_text(color = rev(location_year_name_colors[temp_lyn_order]),
                                     size = 8, hjust = 0.5))

## foi reversible -----
sis_foi_fig <- clu_all %>%
    # restrict to clusters with at least minimum cluster size
    filter(pgp3_n_tested >= min_clu_n) %>% 
    mutate(sis_foi = sis_foi*100) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>% 
    ggplot() +
    geom_point(aes(x = location_year_name, y = sis_foi, color = location_year_name),
               shape = "|", alpha = 0.8, size = 1.5) +
    geom_boxplot(aes(x = location_year_name, y = sis_foi),
                 color = "black", width = 0.5, outlier.shape = NA, fill = NA, lwd = 0.4) +
    scale_x_discrete(limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 70, by = 10), limits = c(0,75)) +
    labs(x = NULL, y = "SCR per 100 PY\n(SRR=6.6)") +
    coord_flip() +
    temp_theme +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 7),
          axis.title.x = element_text(size = 8),
          axis.text.y = element_blank())

## pcr ------
cc_pcr_manual <- data.frame(location_year_name = c("Andabet, Ethiopia 2017",
                                                   "Dera, Ethiopia 2017",
                                                   "Woreta town, Ethiopia 2017",
                                                   "Alefa, Ethiopia 2017"),
                            prev = c(11.3, 0, 0, 0),
                            stringsAsFactors = FALSE)

pcr_fig <- clu_all %>%
    # restrict to clusters with at least minimum cluster size
    filter(pcr_n_tested >= min_clu_n) %>% 
    mutate(location_year_name = factor(location_year_name, levels = rev(temp_lyn_order))) %>% 
    ggplot() +
    geom_point(aes(x = location_year_name, y = pcr_prev, color = location_year_name),
               shape = "|", alpha = 0.8, size = 1.5) +
    geom_boxplot(aes(x = location_year_name, y = pcr_prev),
                 color = "black", width = 0.5, outlier.shape = NA, fill = NA, lwd = 0.4) +
    geom_point(data = cc_pcr_manual %>% filter(location_year_name %in% temp_lyn_order),
             aes(x = location_year_name, y = prev, fill = location_year_name),
             alpha = 0.7, pch = 21, color = "black", size = 2) +
  # add NA labels for studies without PCR measurements
  annotate("text", x= c(1,2,5),y=1, label="NA", size=2) +
    scale_x_discrete(limits = rev(temp_lyn_order)) +
    scale_y_continuous(breaks = seq(0, 70, by = 10)) +
    labs(x = NULL, y = "PCR prevalence (%)") +
    coord_flip(ylim = c(0,72))+
    temp_theme +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 7),
          axis.title.x = element_text(size = 8),
          axis.text.y = element_blank())#,
          #plot.margin = margin(0.2,0.2,0.5,0.2,unit = "cm"))

```

```{r, fig.width = 8.5, fig.height = 3.7}
# combine and save figs
combined_fig <- plot_grid(age_sero_fig, sir_foi_fig, pcr_fig, nrow = 1, rel_widths = c(1.2,1,0.55))

combined_fig_wsub <- add_sub(
  plot = combined_fig,
  label = expression(paste(italic("Abbreviations: "), "PY = person-years; ETH = Ethiopia; NER = Niger; TZA = Tanzania; MWI = Malawi; MAR = Morocco")),
  hjust = 0,
  x = 0.02,
  size = 8
)

ggdraw(combined_fig_wsub)

if(save_figs) {
  ggsave(filename = here("output", "fig2_age_sero_curves.png"),
         ggdraw(combined_fig_wsub),
         width = 8.5, height = 3.7, units = "in")
}

```

## For text

Across populations, estimated seroconversion rates ranged from `r round(min(lyn_all$sir_foi_median, na.rm = TRUE)*100,1)` to `r round(max(lyn_all$sir_foi_median, na.rm = TRUE)*100,1)` per 100 person-years (median: `r round(median(lyn_all$sir_foi_median, na.rm = TRUE)*100,1)`) using a catalytic model without seroreversion.

## Session Info
```{r session info}
sessionInfo()
```
