---
title: "Monitoring transmission intensity of trachoma with serology"
subtitle: "Identifying clusters with infection using serology thresholds"
author: "Contributors: Christine Tedijanto christine.tedijanto@ucsf.edu and Ben Arnold ben.arnold@ucsf.edu"
date: "Last updated: `r Sys.Date()`"
output:
    html_document:
      code_folding: hide
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
number_sections: false
---


## Setup

```{r setup}
#------------------------------
# Load project config file
#------------------------------
library(here)
source(here("R/0-config.R"))

#------------------------------
# global variables
#------------------------------

# minimum number of measurements per cluster
# see 1-trachoma-sero-transmission-make-datasets.Rmd
# for examination of minimum cluster size. 15 chosen
min_clu_n <- 15 

# save figures?
save_figs <- TRUE 

```

## Load saved estimates

```{r load data}
#------------------------------
# load cluster-level estimates
# these estimates were generated
# in the script 2-estimate-foi.Rmd
#------------------------------
clu_all  <- read_rds(here("output", "clu_all_summary.rds")) 
clu_1to5 <- read_rds(here("output", "clu_1to5_summary.rds")) 

#------------------------------
# load study-level estimates
# these estimates were generated
# in the script 2-estimate-foi.Rmd
#------------------------------
lyn_all <- read_rds(here("output", "lyn_all_summary.rds"))

```

```{r}
get_sens_spec <- function(df, dvar, dcutoff, var, range, step){
  
   df <- df %>% filter(!is.na(get(dvar)), !is.na(get(var)))

  # number of D+ "disease-positives" (more infections than cutoff)
  dpos <- df %>% filter(get(dvar)>dcutoff) %>% nrow()
  # number of D- "disease-negatives"
  dneg <- df %>% filter(get(dvar)<=dcutoff) %>% nrow()
  
  temp_cutoffs <- seq(0, range, by = step)
  
  # numerator of sensitivity - T+ & D+
  sens_n <- sapply(temp_cutoffs,
                   function(x) df %>%
                     filter(get(dvar)>dcutoff,
                            get(var)>x) %>%
                     nrow())
  
  # numerator of specificity - T- & D-
  spec_n <- sapply(temp_cutoffs,
                   function(x) df %>%
                     filter(get(dvar)<=dcutoff,
                            get(var)<=x) %>%
                     nrow())
  
  # number of test positives
  tpos <- sapply(temp_cutoffs,
                   function(x) df %>%
                     filter(get(var)>x) %>%
                     nrow())
  
  ret <- data.frame(dvar = dvar, dcutoff = dcutoff,
                    var = var, cutoff = temp_cutoffs,
                    dpos = dpos, sens = sens_n / dpos,
                    dneg = dneg, spec = spec_n / dneg,
                    tpos = tpos, ppv = sens_n / tpos) %>% 
    mutate(spec_comp = 1 - spec,
           youdenj = sens + spec -1,
           f1score = 2*ppv*sens / (ppv + sens))
  
  return(ret)
}
```

```{r}
lyn_all_list <- c("Wag Hemra, Ethiopia (WUHA) 2019", "Wag Hemra, Ethiopia (TAITU) 2018",
                   "Kongwa, Tanzania 2018", "Kongwa, Tanzania 2013",
                   "Chikwawa, Malawi 2014", "Mchinji, Malawi 2014", "Matameye, Niger (PRET) 2013")

lyn_1to5_list <- c("Wag Hemra, Ethiopia (WUHA) 2019", "Wag Hemra, Ethiopia (TAITU) 2018",
                   "Kongwa, Tanzania 2018", 
                   "Chikwawa, Malawi 2014", "Mchinji, Malawi 2014", "Matameye, Niger (PRET) 2013")
```

```{r}
n_steps <- 200
sens_spec_vars <- data.frame(dvar = c("pcr_prev", "pcr_prev"),
                             var = c("pgp3_prev", "sir_foi"),
                             range = c(100, 0.5)) %>%
  mutate(step = range / n_steps) %>% 
  crossing(data.frame(dcutoff = c(0,5)))

sens_spec_all <- lapply(1:nrow(sens_spec_vars),
                       function(x){
                         get_sens_spec(
                           df = clu_all %>%
                             filter(location_year_name %in% lyn_all_list,
                                    pcr_n_tested >= min_clu_n),
                           dvar = sens_spec_vars$dvar[x],
                           dcutoff = sens_spec_vars$dcutoff[x],
                           var = sens_spec_vars$var[x],
                           range = sens_spec_vars$range[x],
                           step = sens_spec_vars$step[x]
                          )}) %>% bind_rows()


sens_spec_all_bylyn <- foreach(curr_lyn = lyn_all_list, .combine = rbind) %dopar% {
  lapply(1:nrow(sens_spec_vars),
                       function(x){
                         get_sens_spec(
                           df = clu_all %>%
                             filter(location_year_name == curr_lyn,
                                    pcr_n_tested>= min_clu_n),
                           dvar = sens_spec_vars$dvar[x],
                           dcutoff = sens_spec_vars$dcutoff[x],
                           var = sens_spec_vars$var[x],
                           range = sens_spec_vars$range[x],
                           step = sens_spec_vars$step[x]
                          )}) %>% bind_rows() %>% mutate(location_year_name = curr_lyn)
}
```

```{r}
sens_spec_1to5 <- lapply(1:nrow(sens_spec_vars),
                       function(x){
                         get_sens_spec(
                           df = clu_1to5 %>%
                             filter(location_year_name %in% lyn_1to5_list,
                                    pcr_n_tested >= min_clu_n),
                           dvar = sens_spec_vars$dvar[x],
                           dcutoff = sens_spec_vars$dcutoff[x],
                           var = sens_spec_vars$var[x],
                           range = sens_spec_vars$range[x],
                           step = sens_spec_vars$step[x]
                          )}) %>% bind_rows()

sens_spec_1to5_bylyn <- foreach(curr_lyn = lyn_1to5_list, .combine = rbind) %dopar% {
  lapply(1:nrow(sens_spec_vars),
                       function(x){
                         get_sens_spec(
                           df = clu_1to5 %>%
                             filter(location_year_name == curr_lyn,
                                    pcr_n_tested >= min_clu_n),
                           dvar = sens_spec_vars$dvar[x],
                           dcutoff = sens_spec_vars$dcutoff[x],
                           var = sens_spec_vars$var[x],
                           range = sens_spec_vars$range[x],
                           step = sens_spec_vars$step[x]
                          )}) %>% bind_rows() %>% mutate(location_year_name = curr_lyn)
}
```

```{r}
get_sens_spec_fig <- function(df, lyn_list, dvar_input, cutoff_var, cutoff_lab){
  
  lyn_temp <- lyn_all %>%
    filter(location_year_name %in% get(lyn_list)) %>% 
    mutate(elim = ifelse(get(paste0(dvar_input, "_median"))==0, 1, 0)) %>% 
    dplyr::select(location_year_name, val = eval(paste0(cutoff_var, "_median")), elim)
  
  temp_df <- df %>% 
    filter(var == cutoff_var,
           dvar == dvar_input,
           dcutoff == 0)
  
  if(cutoff_var %in% c("sir_foi", "sis_foi")){
    temp_df <- temp_df %>% mutate(cutoff = cutoff*100)
    sens_spec_all_bylyn <- sens_spec_all_bylyn %>% mutate(cutoff = cutoff*100)
    lyn_temp <- lyn_temp %>% mutate(val = val*100)
  }
  
  cutoff_90 <- temp_df %>% filter(sens>=0.9) %>% pull(cutoff) %>% max()
  cutoff_80 <- temp_df %>% filter(sens>=0.8) %>% pull(cutoff) %>% max()
  
  temp_df %>% 
    ggplot() +
    # thresholds
    geom_segment(x = cutoff_90, xend = cutoff_90, y = 0, yend = 1, lty = "solid", lwd = 0.7) +
    geom_segment(x = cutoff_80, xend = cutoff_80, y = 0, yend = 1, lty = "dotted", lwd = 0.7) +
    geom_line(data = sens_spec_all_bylyn %>%
                filter(location_year_name %in% get(lyn_list),
                       var == cutoff_var,
                       dvar == dvar_input,
                       dcutoff == 0),
              aes(x = cutoff, y = sens, group = location_year_name, color = location_year_name), lwd = 0.4, alpha = 0.8) +
    geom_line(aes(x = cutoff, y = sens), lwd = 1, color = "black") +
    geom_line(aes(x = cutoff, y = spec), lwd = 1, color = "black") +
    # population points
    geom_point(data = lyn_temp, aes(x = val, y = -0.05, fill = location_year_name),
              pch = 21, size = 2.5, alpha = 0.8) +
    geom_text(aes(x = max(cutoff)*0.9, y = max(spec)*0.95), label = "specificity", color = "black",fontface="bold") +
    geom_text(aes(x = max(cutoff)*0.9, y = min(sens)+0.1), label = "sensitivity", color = "black",fontface="bold") +
    scale_fill_manual(values = location_year_name_colors) +
    scale_color_manual(values = location_year_name_colors) +
    labs(x = cutoff_lab) +
    theme_minimal() +
    theme(axis.title.y = element_blank(),
          legend.position = "none",
          axis.title.x = element_text(size = 10),
          axis.text = element_text(size = 8),
          panel.grid.minor = element_blank(),
          plot.margin = margin(t=16.5,r=5.5,b=5.5,l=20,unit="pt"))
  
}
```

```{r}
sp_fig <- get_sens_spec_fig(df = sens_spec_all,
                            lyn_list = "lyn_all_list",
                            dvar_input = "pcr_prev",
                            cutoff_var = "pgp3_prev",
                            cutoff_lab = "Seroprevalence threshold (%)") 

sir_fig <- get_sens_spec_fig(df = sens_spec_all,
                            lyn_list = "lyn_all_list",
                            dvar_input = "pcr_prev",
                            cutoff_var = "sir_foi",
                            cutoff_lab = "Seroconversion rate per 100 PY threshold") 

sp_fig_1to5 <- get_sens_spec_fig(df = sens_spec_1to5,
                            lyn_list = "lyn_1to5_list",
                            dvar_input = "pcr_prev",
                            cutoff_var = "pgp3_prev",
                            cutoff_lab = "Seroprevalence threshold (%)") 

sir_fig_1to5 <- get_sens_spec_fig(df = sens_spec_1to5,
                            lyn_list = "lyn_1to5_list",
                            dvar_input = "pcr_prev",
                            cutoff_var = "sir_foi",
                            cutoff_lab = "Seroconversion rate per 100 PY threshold") 

```

## ROC curves

```{r}
# calculate AUC
auc_response <- clu_all %>% filter(location_year_name %in% lyn_all_list, pcr_n_tested>= min_clu_n) %>% mutate(pcr_any=as.numeric(pcr_prev>0)) %>% pull(pcr_any)
auc_pred_pgp3 <- clu_all %>% filter(location_year_name %in% lyn_all_list, pcr_n_tested>= min_clu_n) %>% pull(pgp3_prev)
auc_pred_sir <- clu_all %>% filter(location_year_name %in% lyn_all_list, pcr_n_tested>= min_clu_n) %>% pull(sir_foi)
auc_pgp3 <- pROC::auc(auc_response, auc_pred_pgp3)
auc_sir <- pROC::auc(auc_response, auc_pred_sir)

print(paste("AUC all, seroprevalence for PCR>0:", round(auc_pgp3,2)))
print(paste("AUC all, seroconversion for PCR>0:", round(auc_sir,2)))

# ROC curve
temp_df <- sens_spec_all %>% 
    filter(var == "pgp3_prev",
           dvar == "pcr_prev",
           dcutoff == 0)
  
cutoff_90 <- temp_df %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% dplyr::select(sens, spec)
cutoff_80 <- temp_df %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% dplyr::select(sens, spec)

roc_sp_fig <- sens_spec_all %>%
  filter(dvar == "pcr_prev", var == "pgp3_prev", dcutoff==0) %>%
  arrange((1-spec), sens) %>%
  ggplot(aes(x = 1-spec, y = sens)) +
  # add lines for each study pop
  geom_line(data = sens_spec_all_bylyn %>%
              filter(dvar == "pcr_prev", var == "pgp3_prev", dcutoff==0) %>%
              arrange((1-spec), sens),
            aes(x = 1-spec, y = sens, color = location_year_name),
            lwd = 0.4, alpha = 0.8) +
  geom_line(lwd = 1) +
  scale_color_manual(values = location_year_name_colors) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  geom_text(data = data.frame(x_text = 0.75, y_text = 0.95),
            aes(x = x_text, y = y_text), size = 3.5, hjust=0.5, color="black", fontface="bold",
            label = paste0("AUC=", sprintf("%1.2f",auc_pgp3))) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.margin = margin(t=16.5,r=5.5,b=5.5,l=20,unit="pt")
        )

roc_sir_fig <- sens_spec_all %>%
  filter(dvar == "pcr_prev", var == "sir_foi", dcutoff==0) %>%
  arrange((1-spec), sens) %>%
  ggplot(aes(x = 1-spec, y = sens)) +
  # add lines for each study pop
  geom_line(data = sens_spec_all_bylyn %>%
              filter(dvar == "pcr_prev", var == "sir_foi", dcutoff==0) %>%
              arrange((1-spec), sens),
            aes(x = 1-spec, y = sens, color = location_year_name),
            lwd = 0.4, alpha = 0.8) +
  geom_line(lwd = 1) +
  scale_color_manual(values = location_year_name_colors) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  geom_text(data = data.frame(x_text = 0.8, y_text = 0.95),
            aes(x = x_text, y = y_text), size = 3.5, hjust=0.5, color="black",fontface="bold",
            label = paste0("AUC=", sprintf("%1.2f",auc_sir))) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.margin = margin(t=16.5,r=5.5,b=5.5,l=20,unit="pt")
        )

# 1 to 5
# calculate AUC
auc_response <- clu_1to5 %>% filter(location_year_name %in% lyn_1to5_list, pcr_n_tested>= min_clu_n) %>% mutate(pcr_any=as.numeric(pcr_prev>0)) %>% pull(pcr_any)
auc_pred_pgp3 <- clu_1to5 %>% filter(location_year_name %in% lyn_1to5_list, pcr_n_tested>= min_clu_n) %>% pull(pgp3_prev)
auc_pred_sir <- clu_1to5 %>% filter(location_year_name %in% lyn_1to5_list, pcr_n_tested>= min_clu_n) %>% pull(sir_foi)
auc_pgp3 <- pROC::auc(auc_response, auc_pred_pgp3)
auc_sir <- pROC::auc(auc_response, auc_pred_sir)

print(paste("AUC 1-5yo, seroprevalence for PCR>0:", round(auc_pgp3,2)))
print(paste("AUC 1-5yo, seroconversion for PCR>0:", round(auc_sir,2)))

# ROC curve
temp_df <- sens_spec_1to5 %>% 
    filter(var == "pgp3_prev",
           dvar == "pcr_prev",
           dcutoff == 0)
  
cutoff_90 <- temp_df %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% dplyr::select(sens, spec)
cutoff_80 <- temp_df %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% dplyr::select(sens, spec)

roc_sp_1to5_fig <- sens_spec_1to5 %>%
  filter(dvar == "pcr_prev", var == "pgp3_prev", dcutoff==0) %>%
  arrange((1-spec), sens) %>%
  ggplot(aes(x = 1-spec, y = sens)) +
  # add lines for each study pop
  geom_line(data = sens_spec_1to5_bylyn %>%
              filter(dvar == "pcr_prev", var == "pgp3_prev", dcutoff==0) %>%
              arrange((1-spec), sens),
            aes(x = 1-spec, y = sens, color = location_year_name),
            lwd = 0.4, alpha = 0.8) +
  geom_line(lwd = 1) +
  scale_color_manual(values = location_year_name_colors) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  geom_text(data = data.frame(x_text = 0.75, y_text = 0.95),
            aes(x = x_text, y = y_text), size = 3.5, hjust=0.5, color="black", fontface="bold",
            label = paste0("AUC=", sprintf("%1.2f",auc_pgp3))) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.margin = margin(t=16.5,r=5.5,b=5.5,l=20,unit="pt")
        )

roc_sir_1to5_fig <- sens_spec_1to5 %>%
  filter(dvar == "pcr_prev", var == "sir_foi", dcutoff==0) %>%
  arrange((1-spec), sens) %>%
  ggplot(aes(x = 1-spec, y = sens)) +
  # add lines for each study pop
  geom_line(data = sens_spec_all_bylyn %>%
              filter(dvar == "pcr_prev", var == "sir_foi", dcutoff==0) %>%
              arrange((1-spec), sens),
            aes(x = 1-spec, y = sens, color = location_year_name),
            lwd = 0.4, alpha = 0.8) +
  geom_line(lwd = 1) +
  scale_color_manual(values = location_year_name_colors) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  geom_text(data = data.frame(x_text = 0.8, y_text = 0.95),
            aes(x = x_text, y = y_text), size = 3.5, hjust=0.5, color="black",fontface="bold",
            label = paste0("AUC=", sprintf("%1.2f",auc_sir))) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.margin = margin(t=16.5,r=5.5,b=5.5,l=20,unit="pt")
        )

```

```{r}
# add colors and short labels to the lyn_all dataset to make the legend
# filter to studies that contributed to this analysis in lyn_all_list
lyn_colors_df <- data.frame(location_year_name = names(location_year_name_colors), color=location_year_name_colors)
lyn_all2 <- lyn_all %>%
  left_join(lyn_colors_df,by="location_year_name") %>%
  left_join(location_year_name_short, by="location_year_name") %>%
  filter(location_year_name %in% lyn_all_list)

my_ggp <- ggplot(data = lyn_all2) +
  geom_point(aes(x = pgp3_prev_median, y = sir_foi_median, fill = location_year_name),
             pch = 21, color = "black", size = 2.5, alpha = 0.8) +
  scale_fill_manual(values = lyn_all2$color,
                    labels = lyn_all2$lyn_short) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(0.8, "cm"),
        legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 2))
step1 <- ggplot_gtable(ggplot_build(my_ggp))
step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
step3 <- step1$grobs[[step2]]
```

## Combine and save figures


**Figure 6. Seroprevalence among children 1-9y**
```{r save fig6, fig.width = 8.5, fig.height = 3.5}
sens_spec_fig <- plot_grid(plot_grid(sp_fig, roc_sp_fig, nrow = 1, labels = 'auto'),
                           step3, rel_heights = c(1, 0.15), ncol = 1)

sens_spec_fig

ggsave(filename = here("output", "fig6_sens_spec_sp.pdf"),
       sens_spec_fig,
       device = pdf,
       width = 180, height = 90, units = "mm")
```

**Figure S5. Seroconversion rate among children 1-9y**

```{r save figs5, fig.width = 8.5, fig.height = 3.5}
sens_spec_fig_scr <- plot_grid(plot_grid(sir_fig, roc_sir_fig, nrow = 1, labels = 'auto'),
                           step3, rel_heights = c(1, 0.15), ncol = 1)

sens_spec_fig_scr

ggsave(filename = here("output", "figS5_sens_spec_scr.pdf"),
       sens_spec_fig_scr,
       device = pdf,
       width = 180, height = 90, units = "mm")
```

**Figure S6. Seroprevalence and seroconversion rate among children 1-5y**

```{r save figs6, fig.width = 8.5, fig.height = 7}
sens_spec_1to5_fig <- plot_grid(
  plot_grid(sp_fig_1to5, roc_sp_1to5_fig, sir_fig_1to5, roc_sir_1to5_fig, nrow = 2, labels = 'auto'),
  step3, rel_heights = c(1, 0.1), ncol = 1)

sens_spec_1to5_fig

ggsave(filename = here("output", "figS6_sens_spec_sp_1to5.pdf"),
       sens_spec_1to5_fig,
       device = pdf,
       width = 180, height = 180, units = "mm")
```

## Save Supplementary Data files 1-4

```{r save tables}
# save tables

# table s2. seroprevalence 1-9 year olds
write_csv(sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% 
            mutate_at(vars(sens,spec), ~round(.,3)) %>% mutate_at(vars(cutoff), ~round(.,2)) %>% 
            rename(seroprev_threshold = cutoff), 
          file = here("output", "Supplementary_Data_1_seroprev_threshold_sens_spec_1to9years.csv"))

# table s3. seroconversion 1-9 year olds
write_csv(sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% 
            mutate(cutoff = cutoff*100) %>% 
            mutate_at(vars(sens,spec), ~round(.,3)) %>% mutate_at(vars(cutoff), ~round(.,2)) %>% 
            rename(scr_threshold = cutoff), 
          file = here("output", "Supplementary_Data_2_scr_threshold_sens_spec_1to9years.csv")) 

# table s4. seroprevalence 1-5 year olds
write_csv(sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% 
            mutate_at(vars(sens,spec), ~round(.,3)) %>% mutate_at(vars(cutoff), ~round(.,2)) %>% 
            rename(seroprev_threshold = cutoff), 
          file = here("output", "Supplementary_Data_3_seroprev_threshold_sens_spec_1to5years.csv"))

# table s5. seroconversion 1-5 year olds
write_csv(sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% 
            mutate(cutoff = cutoff*100) %>% 
            mutate_at(vars(sens,spec), ~round(.,3)) %>% mutate_at(vars(cutoff), ~round(.,2)) %>% 
            rename(scr_threshold = cutoff),
          file = here("output", "Supplementary_Data_4_scr_threshold_sens_spec_1to5years.csv"))
```

## For text

* Clusters with PCR measurements: `r clu_all %>% filter(location_year_name %in% lyn_all_list, pcr_n_tested>= min_clu_n) %>% nrow()`
* Clusters with PCR measurements with at least one infection detected: `r clu_all %>% filter(location_year_name %in% lyn_all_list, pcr_n_tested>= min_clu_n, pcr_n_pos>0) %>% nrow()`

* Seroprevalence threshold for 90% sensitivity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* Seroconversion (no seroreversion) threshold for 90% sensitivity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* Seroprevalence threshold for 80% sensitivity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* Seroconversion (no seroreversion) threshold for 80% sensitivity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`

* 1-5yo Clusters with PCR measurements: `r clu_1to5 %>% filter(location_year_name %in% lyn_1to5_list, pcr_n_tested>= min_clu_n) %>% nrow()`
* 1-5yo Clusters with PCR measurements with at least one infection detected: `r clu_1to5 %>% filter(location_year_name %in% lyn_1to5_list, pcr_n_tested>= min_clu_n, pcr_n_pos>0) %>% nrow()`

* 1-5yo Seroprevalence threshold for 90% sensitivity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* 1-5yo Seroconversion (no seroreversion) threshold for 90% sensitivity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* 1-5yo Seroprevalence threshold for 80% sensitivity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`
* 1-5yo Seroconversion (no seroreversion) threshold for 80% sensitivity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff)`, specificity: `r sens_spec_1to5 %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi") %>% filter(sens>=0.8) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(spec) %>% round(2)`

* Clusters in Dosso, Alefa, and Woreta town above 90% sensitivity threshold: `r clu_all %>% filter(pgp3_n_tested>=min_clu_n, location_year_name %in% c("Dosso, Niger (MORDOR) 2018", "Alefa, Ethiopia 2017", "Woreta town, Ethiopia 2017"), pgp3_prev>(sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev") %>% filter(sens>=0.9) %>% arrange(desc(cutoff)) %>% slice(1) %>% pull(cutoff))) %>% nrow()`

* 0.015 seroconversions: sensitivity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi", cutoff == "0.015") %>% pull(sens) %>% round(3)`, specificity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "sir_foi", cutoff == "0.015") %>% pull(spec) %>% round(3)`
* 6.0% seroprevalence: sensitivity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev", cutoff == "6") %>% pull(sens) %>% round(3)`, specificity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev", cutoff == "6") %>% pull(spec) %>% round(3)`
* 6.5% seroprevalence: sensitivity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev", cutoff == "6.5") %>% pull(sens) %>% round(3)`, specificity = `r sens_spec_all %>% filter(dvar == "pcr_prev", dcutoff == 0, var == "pgp3_prev", cutoff == "6.5") %>% pull(spec) %>% round(3)`


## Session Info
```{r session info}
sessionInfo()
```